---
title: "Culture_amplicons_DADA2"
author: "Chance English"
date: "10/23/2024"
output: html_document
---


```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, echo = FALSE, warning = FALSE)
```


```{r, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}

library(tidyverse)
library(dada2)
library(ShortRead)
library(beepr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyverse)
library(here)
library(ggpubr)


library(janitor)
library(ggsci)
library(lubridate)
library(readxl)
library(growthcurver)
library(zoo)
library(scales)
library(kableExtra)
library(broom)

library(dplyr)
library(RColorBrewer)

library(reshape2)
library(vegan)
library(indicspecies)
library(readr)
library(ggrepel)

```



```{r}

path <- "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Inputdata/fastq_cultures"

#store names of forward and reverse files as lists 

fnFS <- list.files(path, pattern = "_R1_001.fastq.gz", full.names = TRUE)

fnRS <- list.files(path, pattern = "_R2_001.fastq.gz", full.names = TRUE)


```


# Retreive Orientation of Primers 

Our primers target the V4 region and are known as 514F-Y and 806RB primers (Appril et al. 2015)

```{r}

FWD <- "GTGYCAGCMGCCGCGGTAA"
REV <- "GGACTACNVGGGTWTCTAAT"

all_orients <- function(primer) {
  require(Biostrings)
  dna <- DNAString(primer)
  orients <- c(forward = dna, complement = complement(dna), reverse = reverse(dna),
               reverse_complement = reverseComplement(dna))
  return(sapply(orients, toString))
}


# Store Forward and Reverse orientations separately 

FWD.orients <- all_orients(FWD)
REV.orients <- all_orients(REV)




```



# Search for Primers


```{r}

primerHits <- function(primer, fn) {
  nhits <- vcountPattern(primer, sread(readFastq(fn)), fixed = FALSE)
  return(sum(nhits>0))
}


rbind(FWD.ForwardReads = sapply(FWD.orients, primerHits, fn = fnFS[[1]]),
      FWD.ReverseReads = sapply(FWD.orients, primerHits, fn = fnRS[[1]]),
      REV.ForwardReads = sapply(REV.orients, primerHits, fn = fnFS[[1]]),
      REV.ReverseReads = sapply(REV.orients, primerHits, fn = fnRS[[1]]))




```

# Inspect Read Quality Profiles 

Should always look at some of your sequences to assess the quality of the sequencing run. 

## Forward Reads 

```{r, fig.height=10, fig.width=12}

plotQualityProfile(fnFS[24:34])


# Good score, will trim forward at 240

foo1 <- readFastq(fnFS[[24:34]])




```


```{r}

plotQualityProfile(fnRS[1:5])

# For samples 24:34 pretty high quality score to about 220. Running for samples 1:5 also

# Lower quality score for samples 1:5 drops between 150 - 200




```

# Filtering and Trimming

Don't want to trim our raw sequences. We want to create a new folder that contains the new trimmed sequences. 


```{r}

#Get sampole names

#define the basenames of the FnFS as the first part of each file name until the "_L"
#Apply this to all files 


sample.names <- sapply(strsplit(basename(fnFS), "_L00"), '[',1)
sample.names

# Create a filtered folder in the working directory as a palce to put all the new filtered fastq files.

filt_path <- file.path(path, "filtered_culture_amplicons_path")

# Add the appropriate designation string to any new files made that will be put nto the filtered path 

filtFS <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq"))
filtRS <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq"))


```


```{r}

out <- filterAndTrim(fnFS, filtFS, fnRS, filtRS, truncLen = c(240,150), maxN = 0, 
                     maxEE = c(2,2), truncQ = 2, rm.phix = TRUE, compress = TRUE)



out

```


# Learn the error rates


```{r}

errF <- learnErrors(filtFS, multithread = TRUE)

errR <- learnErrors(filtRS, multithread = TRUE)

beep(sound = 8)
```


```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=12, fig.align='center'}


plotErrors(errF, nominalQ = TRUE)





```




```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=12, fig.align='center'}


plotErrors(errR, nominalQ = TRUE)



```


How well does the error model (black line) fit the errors (black dots)? Also the error rate (y axis) should generally decrease with increase quality score. 



# Dereplication 


Dada2 will combine all replicate sequences into once unique sequence, keeping track of the number of identical sequences



```{r}

derepFS <- derepFastq(filtFS, verbose = TRUE)
derepRS <- derepFastq(filtRS, verbose = TRUE)

# Name the derep samples by the sample names 

names(derepFS) <- sample.names
names(derepRS) <- sample.names


```


# Infer the sequence variants 

Apply core dada2 sample inference algorithm to the dereplicated data 

Infer sequence variants in each sample taking out the variants that have excessive error rates. (i.e we are applying the error models to the data). Apply earlier model from a subset of the data to the whole dataset to determine which sequences are real and which are not. 


```{r}

dadaFS <- dada(derepFS, err = errF, multithread = TRUE)
dadaRS <- dada(derepRS, err = errR, multithread = TRUE)
beep(3)


```


Merge the overlapping reads. This will decrease the number of sequence variants. Will decrease the total number of sequences. 


```{r}

mergers <- mergePairs(dadaFS, derepFS, dadaRS, derepRS, verbose = TRUE, trimOverhang = TRUE)

beep(6)

#trimOverhang overhang will remove hits from earlier when checking for primers. 

```




```{r}

head(mergers[[1]])



```

Save the unassigned merger data into output data file


```{r}



saveRDS(mergers, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/dada_merged_cultures.rds")

mergers <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/dada_merged_cultures.rds")

```


Construct a sequence table of our samples that is analagous to the OTU table produced by classical methods

```{r}

seqtable <- makeSequenceTable(mergers)


dim(seqtable)

df_seq_table <- as.data.frame(seqtable, row.names = TRUE)

write.table(t(seqtable), file="cluture_seqtable.txt",sep = ",",row.names=FALSE, col.names=TRUE)

write_csv(df_seq_table,  "/Users/chanceenglish/Desktop/df_seq_table.csv")

```


Check the distribution of the sequence lengths 

```{r}

table(nchar(getSequences(seqtable)))


```

# Remove Chimeras 

In POCR two or more biological sequences can attach to eachother and then polymerase builds a non-biological sequence

```{r}

seqtable_nochim <- removeBimeraDenovo(seqtable, verbose = TRUE)




dim(seqtable_nochim)

```

# Assign Taxonomy using Silva Reference Database (v138)

```{r}

taxa <- assignTaxonomy(seqtable_nochim, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Inputdata/silva_nr99_v138.1_wSpecies_train_set.fa.gz", multithread = TRUE)

beep(10)



```
