---
title: "MP_Biofilm_DADA2"
author: "Chance English"
date: "9/13/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, echo = FALSE, warning = FALSE)
```


```{r}

library(tidyverse)
library(dada2)
library(ShortRead)
library(beepr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyverse)
library(here)
library(ggpubr)


library(janitor)
library(ggsci)
library(lubridate)
library(readxl)
library(growthcurver)
library(zoo)
library(scales)
library(kableExtra)
library(broom)

library(dplyr)
library(RColorBrewer)

library(reshape2)
library(vegan)
library(indicspecies)
library(readr)
library(ggrepel)

```


# Import Files 


```{r}

path <- "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Inputdata/fastq"

#store names of forward and reverse files as lists 

fnFS <- list.files(path, pattern = "_R1_001.fastq", full.names = TRUE)

fnRS <- list.files(path, pattern = "_R2_001.fastq", full.names = TRUE)




```


# Retreive Orientation of Primers 

Our primers target the V4 region and are known as 514F-Y and 806RB primers (Appril et al. 2015)

```{r}

FWD <- "GTGYCAGCMGCCGCGGTAA"
REV <- "GGACTACNVGGGTWTCTAAT"

all_orients <- function(primer) {
  require(Biostrings)
  dna <- DNAString(primer)
  orients <- c(forward = dna, complement = complement(dna), reverse = reverse(dna),
               reverse_complement = reverseComplement(dna))
  return(sapply(orients, toString))
}


# Store Forward and Reverse orientations separately 

FWD.orients <- all_orients(FWD)
REV.orients <- all_orients(REV)




```



# Search for Primers


```{r}

primerHits <- function(primer, fn) {
  nhits <- vcountPattern(primer, sread(readFastq(fn)), fixed = FALSE)
  return(sum(nhits>0))
}


rbind(FWD.ForwardReads = sapply(FWD.orients, primerHits, fn = fnFS[[1]]),
      FWD.ReverseReads = sapply(FWD.orients, primerHits, fn = fnRS[[1]]),
      REV.ForwardReads = sapply(REV.orients, primerHits, fn = fnFS[[1]]),
      REV.ReverseReads = sapply(REV.orients, primerHits, fn = fnRS[[1]]))




```

# Inspect Read Quality Profiles 

Should always look at some of your sequences to assess the quality of the sequencing run. 

## Forward Reads 

```{r, fig.height=10, fig.width=12}

plotQualityProfile(fnFS[24:34])


# Good score, will trim forward at 240


```



```{r}

plotQualityProfile(fnRS[1:5])

# For samples 24:34 pretty high quality score to about 220. Running for samples 1:5 also

# Lower quality score for samples 1:5 drops between 150 - 200




```

# Filtering and Trimming

Don't want to trim our raw sequences. We want to create a new folder that contains the new trimmed sequences. 


```{r}

#Get sampole names

#define the basenames of the FnFS as the first part of each file name until the "_L"
#Apply this to all files 


sample.names <- sapply(strsplit(basename(fnFS), "_L00"), '[',1)
sample.names

# Create a filtered folder in the working directory as a palce to put all the new filtered fastq files.

filt_path <- file.path(path, "filtered_MP_Biofilm_path")

# Add the appropriate designation string to any new files made that will be put nto the filtered path 

filtFS <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq"))
filtRS <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq"))


```


```{r}

out <- filterAndTrim(fnFS, filtFS, fnRS, filtRS, truncLen = c(240,150), maxN = 0, 
                     maxEE = c(2,2), truncQ = 2, rm.phix = TRUE, compress = TRUE)



out

```


# Learn the error rates


```{r}

errF <- learnErrors(filtFS, multithread = TRUE)

errR <- learnErrors(filtRS, multithread = TRUE)

beep(sound = 8)
```




```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=12, fig.align='center'}


plotErrors(errF, nominalQ = TRUE)





```




```{r, warning=FALSE, echo=FALSE, message=FALSE, fig.height=10, fig.width=12, fig.align='center'}


plotErrors(errR, nominalQ = TRUE)



```


How well does the error model (black line) fit the errors (black dots)? Also the error rate (y axis) should generally decrease with increase quality score. 



# Dereplication 


Dada2 will combine all replicate sequences into once unique sequence, keeping track of the number of identical sequences



```{r}

derepFS <- derepFastq(filtFS, verbose = TRUE)
derepRS <- derepFastq(filtRS, verbose = TRUE)

# Name the derep samples by the sample names 

names(derepFS) <- sample.names
names(derepRS) <- sample.names


```


# Infer the sequence variants 

Apply core dada2 sample inference algorithm to the dereplicated data 

Infer sequence variants in each sample taking out the variants that have excessive error rates. (i.e we are applying the error models to the data). Apply earlier model from a subset of the data to the whole dataset to determine which sequences are real and which are not. 


```{r}

dadaFS <- dada(derepFS, err = errF, multithread = TRUE)
dadaRS <- dada(derepRS, err = errR, multithread = TRUE)
beep(3)


```


Merge the overlapping reads. This will decrease the number of sequence variants. Will decrease the total number of sequences. 


```{r}

mergers <- mergePairs(dadaFS, derepFS, dadaRS, derepRS, verbose = TRUE, trimOverhang = TRUE)

beep(6)

#trimOverhang overhang will remove hits from earlier when checking for primers. 

```



```{r}

head(mergers[[1]])



```

Save the unassigned merger data into output data file


```{r}



saveRDS(mergers, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/dada_merged_MP_Biofilm.rds")

mergers <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/dada_merged_MP_Biofilm.rds")

```


Construct a sequence table of our samples that is analagous to the OTU table produced by classical methods

```{r}

seqtable <- makeSequenceTable(mergers)


dim(seqtable)


write.table(t(seqtable), file="seqtable.txt",sep = ",",row.names=FALSE, col.names=TRUE)



```

Check the distribution of the sequence lengths 

```{r}

table(nchar(getSequences(seqtable)))


```

# Remove Chimeras 

In POCR two or more biological sequences can attach to eachother and then polymerase builds a non-biological sequence

```{r}

seqtable_nochim <- removeBimeraDenovo(seqtable, verbose = TRUE)




dim(seqtable_nochim)

```

# Assign Taxonomy using Silva Reference Database (v138)

```{r}

taxa <- assignTaxonomy(seqtable_nochim, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Inputdata/silva_nr99_v138.1_wSpecies_train_set.fa.gz", multithread = TRUE)

beep(10)



```


Create a table out of the taxa data (one with the sequences and assignments, one with just the taxa.)


```{r}




saveRDS(t(seqtable_nochim), "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/seqtable_nochim.rds")

saveRDS(taxa, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/taxa.rds")

beep(6)

```

# Create a Seqtable with Taxa and Sample Names for Columns


```{r}

seqtable_df <- as.data.frame(seqtable)

taxatable_df <- as.data.frame(taxa.table)

sampletable_df <- as.data.frame(sample.table)



seqtable_df_t <- as.data.frame(t(seqtable_df)) 

seqtable_df_t <- seqtable_df_t %>% 
  rownames_to_column(var = "sequence")


taxa_table_sequence_rownames <- taxatable_df %>% 
  rownames_to_column(var = "sequence")


seqtab_nochim_taxa <- merge(seqtable_df_t, taxa_table_sequence_rownames, by = "sequence") %>% 
  as.data.frame()

write.csv(seqtab_nochim_taxa, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/seqtab_nochim_taxa.csv", row.names = FALSE)

```


# NMDS


```{r}

sampletable_df_nmds <- read_csv(here("Inputdata", "sampledata", "MP_Biofilm_Samples_Metadata.csv"))



seqtab_nochim_taxa <- read_csv(here("Outputdata", "seqtab_nochim_taxa.csv"))


MP_Biofilm_counts_taxa_table <- seqtab_nochim_taxa %>% 
  select(sequence, MP_230809_B1_S1:MP_240616_B6_S54, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  filter(Family != "Mitochondria" & Order != "Chloroplast")


# Filter out samples with bad PCR results 

MP_Biofilm_counts_taxa_table_taxa <- MP_Biofilm_counts_taxa_table %>% 
  select(sequence, c(Kingdom:Species))
  

MP_Biofilm_counts_taxa_table_goodPCR <- MP_Biofilm_counts_taxa_table %>% 
  select(-c(Kingdom:Species)) %>% 
  column_to_rownames(var = "sequence") %>% 
  .[,colSums(.) >15000] %>% 
  rownames_to_column(var = "sequence")

# Return dataframe containing only samples with greater than 10,000 sequence reads 

MP_Biofilm_counts_taxa_table_goodPCR_wTaxa <- merge(MP_Biofilm_counts_taxa_table_goodPCR, MP_Biofilm_counts_taxa_table_taxa, by = "sequence")

MP_Biofilm_counts_taxa_table_goodPCR_wTaxa_long <- MP_Biofilm_counts_taxa_table_goodPCR_wTaxa %>% 
  select(-c(Kingdom:Species)) %>% 
  column_to_rownames(var = "sequence") %>% 
  t() 
  


raremax <- min(rowSums(MP_Biofilm_counts_taxa_table_goodPCR_wTaxa_long))

MP_Biofilm_counts_table_rarefied <- rrarefy(MP_Biofilm_counts_taxa_table_goodPCR_wTaxa_long, raremax)

# Transpose to remove sequences that have 0 counts. Residual from other sampled from different envronment and project from the DY111 sample set

MP_Biofilm_counts_table_rarefied_long <- MP_Biofilm_counts_table_rarefied %>% 
  t()

MP_Biofilm_counts_table_rarefied_long.df <- as.data.frame(MP_Biofilm_counts_table_rarefied_long)

MP_Biofilm_counts_table_rarefied_long.df[is.na(MP_Biofilm_counts_table_rarefied_long.df)] <- 0

MP_Biofilm_counts_table_rarefied_long.df[] <- lapply(MP_Biofilm_counts_table_rarefied_long.df, as.numeric)

MP_Biofilm_counts_table_rarefied_long.df <- MP_Biofilm_counts_table_rarefied_long.df %>%
  filter(rowSums(across(where(is.numeric)))!=0)

# Transpose back to calculate NMDS for quality filtered and rarefied DY111 data set 

MP_Biofilm_counts_table_rarefied_long.df <- MP_Biofilm_counts_table_rarefied_long.df %>% 
  t()

com <- MP_Biofilm_counts_table_rarefied_long.df[,1:ncol(MP_Biofilm_counts_table_rarefied_long.df)] 



m_com <- as.matrix(com)

nmds = metaMDS(m_com, dsitance = "bray")



data.scores = as.data.frame(scores(nmds)) %>% 
  rownames_to_column(var = "DNA_SampleID")



```

# Merge NMDS with Sample Metadata

```{r}


MP_Biofilm_Sample_Data_NMDS <- merge(sampletable_df_nmds, data.scores, by = "DNA_SampleID")




# Plot results of community composition across potential density 

NMDS_p_biofilm <- MP_Biofilm_Sample_Data_NMDS %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = age_d, shape = season)) +
  geom_point(size = 7) +
  labs(color = "Age (d)",
       shape = "") +
  scico::scale_color_scico(palette = "vik", breaks = c(0,25,50,75,100)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 




```


# Calculate diversity metrics 


```{r}


# Rarefied Sequence and Counts Table: No Taxa

MP_Biofilm_rarefied_seqtab_nochim_notaxa.df <- as.data.frame(MP_Biofilm_counts_table_rarefied_long)

MP_Biofilm_rarefied_seqtab_nochim_notaxa.mx <- as.matrix(MP_Biofilm_counts_table_rarefied_long)
# Taxa Table for onlt Taxa with Sequences in the Counts Table

MP_Biofilm_rarefied_seqtab_nochim_notaxa_sequence.mx <- as.matrix(MP_Biofilm_rarefied_seqtab_nochim_notaxa.df %>% 
  rownames_to_column(var = "sequence")  %>% 
  select(sequence))


MP_Biofilm_all_taxa <- MP_Biofilm_counts_taxa_table  %>% 
  select(c(sequence,Kingdom:Species))

MP_Biofilm_rarefied_taxa <- as.matrix(merge(MP_Biofilm_rarefied_seqtab_nochim_notaxa_sequence.mx, MP_Biofilm_all_taxa, by = "sequence") %>% 
  column_to_rownames(var = "sequence"))



# Save Rarefied Sequence Counts and Taxa as RDS for phyloseq


saveRDS(MP_Biofilm_rarefied_seqtab_nochim_notaxa.mx,"/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/rarefied_seqtab_notaxa.rds")

saveRDS(MP_Biofilm_rarefied_taxa,"/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/rarefied_taxa.rds")

saveRDS(MP_Biofilm_Sample_Data_NMDS,"/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sample_table.rds")

# Read in RDS Count and Taxa Table

count_table_diversity <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/rarefied_seqtab_notaxa.rds")

taxa_table_diversity <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/rarefied_taxa.rds")

sample.table_diversity <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sample_table.rds") %>% 
  drop_na(DNA_SampleID) %>% 
  column_to_rownames(var = "DNA_SampleID")


# Generate Phyloseq Object

OTU = otu_table(count_table_diversity, taxa_are_rows = TRUE)
TAX = tax_table(taxa_table_diversity)


ps = phyloseq(OTU, TAX)


# Calcualte Richness Statistics 

richness_df <- estimate_richness(ps, measures = c("Chao1", "Shannon")) %>%
  rownames_to_column(., var = "DNA_SampleID")


MP_biofilm_samples_richness <- merge(MP_Biofilm_Sample_Data_NMDS, richness_df, by = "DNA_SampleID")





# Plot Chao1

 Shannon_div_plot <- MP_biofilm_samples_richness%>%
  ggplot(aes(x = age_d, y = Shannon, color = season)) +
  geom_point() +
  labs(x = "Age (d)",
       y = "Shannon (H')",
       color = "") +
   stat_smooth(aes(color = season), method = "lm", formula = y ~ poly(x, 2), size = 2, se = FALSE) +
   scale_color_manual(values = c("red", "black")) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.position = c(0.8, 0.25),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 

```


Use rarefied data to calcualte Relative Abundances 


```{r}


count.table <- as.data.frame(count_table_diversity) %>% 
  rownames_to_column(var = "sequence")
  
taxa.table <- as.data.frame(taxa_table_diversity) %>% 
  rownames_to_column(var = "sequence")
  


seq_table_taxa_order <- merge(count.table, taxa.table, by = "sequence") %>% 
  mutate(kpcofgs = paste(Kingdom,Phylum,Class,Order,  Family, Genus, Species, sep = "_")) %>% 
  select(-c(sequence, Kingdom:Species)) %>% 
  mutate(rown_num = seq(1:4355), kpcofgs_number = paste(kpcofgs, rown_num, sep = ";")) %>% 
  column_to_rownames(var = "kpcofgs_number") %>% 
  select(-c(kpcofgs, rown_num)) %>% 
  mutate(total_counts = rowSums(.)) %>% 
  filter(total_counts != 0)


  
asv_id <- seq_table_taxa_order %>%   
  arrange(-total_counts) %>% 
  rownames_to_column(var = "kpcofgs_number") %>% 
  mutate(number = seq(1:2891), ASV = "ASV", asv_id = paste(ASV,number, sep = "")) %>% 
  select(asv_id, kpcofgs_number)

seq_table_taxa_order_1 <- seq_table_taxa_order %>% 
  rownames_to_column(var = "kpcofgs_number")



ASV_taxa_number_counts <- merge(seq_table_taxa_order_1, asv_id, by = "kpcofgs_number")
  
  
ASV_taxa_number_RA <- ASV_taxa_number_counts %>%  
  separate(kpcofgs_number, into = c("kpcofgs", "number"), sep = ";") %>% 
  select(-c(total_counts, number)) %>% 
  mutate(asv_rank_taxa = paste(asv_id,kpcofgs, sep = "_")) %>% 
  column_to_rownames(var = "asv_rank_taxa") %>% 
  select(-c(asv_id, kpcofgs)) %>% 
  mutate(across(where(is.numeric), ~ ./sum(.)))
  

seq_table_taxa_RA_long <-as.data.frame(t(ASV_taxa_number_RA))

seq_table_taxa_RA_long_filt <- seq_table_taxa_RA_long  %>% 
  .[,colSums(.) >0]


seq_table_taxa_RA_long_filt.df <- as.data.frame(seq_table_taxa_RA_long_filt)

seq_table_taxa_RA_long_filt.df<- seq_table_taxa_RA_long_filt.df %>% 
  rownames_to_column(var = "DNA_SampleID")


MP_Biofilm_Samples_Diversity_RA <- merge(MP_biofilm_samples_richness, seq_table_taxa_RA_long_filt.df, by = "DNA_SampleID")

write_csv(MP_Biofilm_Samples_Diversity_RA, )

```



# Plot Specific ASV's over time 

```{r}


MP_Biofilm_Samples_Diversity_RA %>% 
  ggplot(aes(x = age_d, y = ASV10_Bacteria_Proteobacteria_Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae_Granulosicoccus_NA)) +
  geom_point(size = 5) +
  geom_smooth(aes(color = season), size = 1, se = FALSE) +
  theme(panel.grid.major = element_line(color = "grey", size = 0.5),
        panel.grid.minor = element_line(color = "grey", size = 0.5),
        strip.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        strip.text.x = element_text(),
        legend.key = element_rect(fill = "white"),
        legend.title = element_text(size = ),
        legend.text = element_text(size = 15),
        strip.text = element_text(size = "20"),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 17),
        text = element_text(family = "Times"))# +
    labs(x = "ASV39_Verrucomicrobiales (% Relative Abundance)",
       y = "Depth (m)") 







```


