---
title: "Sugars"
author: "Chance English"
date: "10/3/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```


```{r}

library(tidyverse)
library(dada2)
library(ShortRead)
library(beepr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyverse)
library(here)
library(ggpubr)
library(janitor)
library(ggsci)
library(lubridate)
library(readxl)
library(growthcurver)
library(zoo)
library(scales)
library(kableExtra)
library(broom)
library(reshape2)
library(vegan)
library(indicspecies)
library(readr)
library(ggrepel)
library(DESeq2)
library(factoextra)

library(cowplot)

```



# Sugar Monomers Composition & PCA


```{r}


sugars <- read_csv(here("Outputdata", "MP_DOM_sugar.csv"))



sugars_subset <- sugars %>% 
  select(DNA_SampleID, Fucose:`Man-URA`) %>% 
  column_to_rownames(var = "DNA_SampleID")




suga_numeric <- sugars[1:42,7:15]

MP_DOM_Sugar_mature <- MP_DOM_Sugar %>% 
  filter(age_d < 20) 


suga_mature_numeric <- MP_DOM_Sugar[1:12, 6:13]

```


```{r}



sug.pca1 <- prcomp(suga_numeric, scale = TRUE)

fviz_eig(sug.pca1)

sug.pca1

summary(sug.pca1)

get_eig(sug.pca1)

eigenvals(sug.pca1)

print(sug.pca1)

# Mature Kelp

sug.pca_mature <- prcomp(suga_mature_numeric, scale = TRUE)

fviz_eig(sug.pca_mature)

sug.pca_mature

summary(sug.pca_mature)

get_eig(sug.pca_mature)

eigenvals(sug.pca_mature)

```


# Sugars Biplot for Mature vs. Senescent Kelp Exudates


```{r}

fviz_pca_biplot(sug.pca1,
             col.var = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             col.ind = "black",
             repel = TRUE     # Avoid text overlapping
             )


pca_plot <- fviz_pca_biplot(sug.pca1, 
                level = "var",
                habillage = sugars[1:42,]$phase, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("#266A25","#BD9C3D")) +
  labs(title = "", 
       x = "PC1 (24%)",
       y = "PC2 (24%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = c(0.2,0.2),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
# PCA Plot Season 

pca_plot_season <- fviz_pca_biplot(sug.pca1, 
                level = "var",
                habillage = MP_DOM_Sugar[1:42,]$season, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("#022E73", "#9D3709")) +
  labs(title = "", 
       x = "PC1 (26%)",
       y = "PC2 (23%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = c(0.2,0.1),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
  
# Compare Mature Kelp for Summer and Spring

fviz_pca_biplot(sug.pca_mature,
             col.var = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             col.ind = "black",
             repel = TRUE     # Avoid text overlapping
             )


pca_plot_mature <- fviz_pca_biplot(sug.pca_mature, 
                level = "var",
                habillage = MP_DOM_Sugar_mature[1:12,]$season, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("steelblue1", "firebrick1")) +
  labs(title = "", 
       x = "PC1 (26%)",
       y = "PC2 (23%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = c(0.15,0.10),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
  





```


# Permanova for Sugar Mol% between Mature and Senescent phase Kelp


```{r}

MP_DOM_sugar_c <- scale(sugars[1:42,7:15])

# PERMANOVA of differences in carbohydrate composition by physiological state

MP_DOM_sugar_physstate_adonis <- adonis(MP_DOM_sugar_c~phase, data = sugars[1:42,], method = "eu")

MP_DOM_sugar_physstate_adonis # R2 = 0.14 p = 0.001


# PERMANOVA of differences in carbohydrate composition by season 

MP_DOM_sugar_season_adonis <- adonis(MP_DOM_sugar_c~season, data = sugars[1:42,], method = "eu")

MP_DOM_sugar_season_adonis # R2 = 0.07 p = 0.007 **

MP_DOM_sugar_phys_state_season_adonis <- adonis(MP_DOM_sugar_c~ phase + season, data = MP_DOM_Sugar[1:42,], method = "eu")

MP_DOM_sugar_phys_state_season_adonis

pairwiseAdonis::pairwise.adonis2(MP_DOM_sugar_phys_state_season_adonis)

```


# Fucose & Mannuronic Acid Boxplot over Time: SUMMER


```{r}

fucose_manURA_age <- MP_DOM_Sugar %>% 
  filter(season == "Summer") %>% 
  select(age_d, Fucose, `Man-URA`, blade) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent") %>% 
  mutate(age_d = as.factor(age_d))
  
boxplot_fucose_manURA_summer <- fucose_manURA_age %>% 
  ggplot(aes(x = age_d, y = mol_percent*100, fill = name)) +
  geom_boxplot() +
   ggtitle(label = "Summer") + 
  labs(x = "Age (d)",
       y = "Mol%",
       fill = "") +
  scale_fill_manual(values = c("steelblue1", "firebrick1")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        plot.title = element_text(size = 24),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.position = "none",
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))




```


# Fucose & Mannuronic Acid Boxplot over Time: SPRING


```{r}

fucose_manURA_age_spring <- MP_DOM_Sugar %>% 
  filter(season == "Spring") %>% 
  select(age_d, Fucose, `Man-URA`, blade) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent") %>% 
  mutate(age_d = as.factor(age_d))
  
boxplot_fucose_manURA_spring <- fucose_manURA_age_spring %>% 
  ggplot(aes(x = age_d, y = mol_percent*100, fill = name)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Mol%",
       fill = "") +
  ggtitle(label = "Spring") + 
  scale_fill_manual(values = c("steelblue1", "firebrick1")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        plot.title = element_text(size = 24), 
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))




```

# Summary: Avg Sugar mol% in Mature vs. Senescent phase

```{r}

sugar_summary <- MP_DOM_Sugar %>% 
  select(phase, season,Fucose:`Man-URA`) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent")%>% 
  group_by(phase, name, season) %>% 
  summarize(mean_molp = mean(mol_percent),
            sd_molp = sd(mol_percent))
  
sugars_fold_change_phys_state <- sugar_summary %>% 
  group_by(name) %>% 
  summarise(fold_change = mean_molp[phase == "Senescent"]/mean_molp[phase == "Mature"])

sugars_fold_change_season <- sugar_summary %>% 
  group_by(phase, name) %>% 
  summarise(fold_change = mean_molp[season == "Spring"]/mean_molp[season == "Summer"])

```


# Glucosamine Mature (16 days) Box Plot between Spring and Summer


```{r}

glucosamine_season_mature <- MP_DOM_Sugar_mature %>% 
  select(season, Glucoseamine, blade) %>% 
  pivot_longer(cols = Glucoseamine,
               names_prefix = "sugar",
               values_to = "mol_percent") 
  
boxplot_glucosamine_season_mature <- glucosamine_season_mature %>% 
  ggplot(aes(x = season, y = mol_percent*100)) +
  geom_boxplot(fill = "grey") +
  labs(x = "",
       y = "Mol%",
       fill = "") +
  scale_fill_manual(values = c("grey")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))



```

# Glucosamine age and season 

```{r}

MP_DOM_Sugar_glucosamine <- sugars %>% 
  filter(age_d %in% c(16, 77, 78)) %>% 
  select(age_d, season, blade, Glucoseamine) %>% 
  mutate(phase = case_when(
    age_d >= 50 ~ "Senescent",
    age_d < 50 ~ "Mature"
  ))


glucosamine_season <- MP_DOM_Sugar_glucosamine %>% 
  select(age_d, season, Glucoseamine, blade, phase) %>% 
  pivot_longer(cols = Glucoseamine,
               names_prefix = "sugar",
               values_to = "mol_percent") 


glucosamine_molp_boxplot <- glucosamine_season %>% 
  ggplot() +
  geom_boxplot(aes(x = season, y = mol_percent*100, fill = phase)) +
  geom_point(aes(x = season, y = mol_percent*100, color = phase),
             position = position_jitterdodge(0.1)) +
  labs(y = "Glucosamine (mole%)",
       x = "",
       color = "",
       fill = "") +
  scale_fill_manual(values = c("#266A25","#BD9C3D")) +
  scale_color_manual(values = c("darkgreen","darkgoldenrod")) +
  scale_y_continuous(limits = c(0,20)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = c(0.72,0.75),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

# Two-way ANOVA Glucosamine vs. Season and Physiology


aov.glucosamine.season.phase <- aov(data = MP_DOM_Sugar_glucosamine, Glucoseamine ~ season * phase)

summary(aov.glucosamine.season.phase)

TukeyHSD(aov.glucosamine.season.phase)


```


# Gluconic acid age and season 


```{r}

MP_DOM_Sugar_glcURA <- MP_DOM_Sugar %>% 
  filter(age_d %in% c(16, 77, 78)) %>% 
  select(age_d, season, blade, `Glc-URA`) 


glcURA_season <- MP_DOM_Sugar_glcURA %>% 
  select(age_d, season, `Glc-URA`, blade) %>% 
  pivot_longer(cols = 'Glc-URA' ,
               names_prefix = "sugar",
               values_to = "mol_percent") 


glcURA_molp_boxplot <- glcURA_season %>% 
  mutate(phase = case_when(age_d < 50 ~ "Mature",
                           age_d > 50 ~ "Senescent")) %>% 
   ggplot(aes(phase, mol_percent)) +
  geom_boxplot(position = "dodge") +
  geom_point(size = 3) +
  facet_wrap(~ season, strip.position = "bottom") +
  theme_minimal() +
   theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.placement = "outside",
        strip.background = element_rect(fill = "white", color = "white"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"),
        strip.text = element_text(size = 20, face = "bold")) +
  labs(x = "",
       y = "Glc-URA (Mol%)")



```

