---
title: "MP_Biofilm_Amplicon_Analysis"
author: "Chance English"
date: "9/16/2024"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```


```{r,include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}

library(tidyverse)
library(dada2)
library(ShortRead)
library(beepr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(tidyverse)
library(here)
library(ggpubr)
library(janitor)
library(ggsci)
library(lubridate)
library(readxl)
library(growthcurver)
library(zoo)
library(scales)
library(kableExtra)
library(broom)
library(reshape2)
library(vegan)
library(indicspecies)
library(readr)
library(ggrepel)
library(DESeq2)

#library(cowplot)

```



# Load in sequence table with taxa 



```{r, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}

# Load Sequence table with sequences for rownames and sample ID for column names. Populated with read number

seq_table <- read_rds("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/seqtable_nochim.rds")

sample_data <- read_rds(as.matrix("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sample.data.rds", header = T, row.names = F, check.names = F))


count_table <- read_csv(as.matrix("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/seqtab_nochim_taxa.csv", header = T, row.names = 1, check.names = F))


sampletable_df_nmds <- read_csv(here("Inputdata", "sampledata", "MP_Biofilm_Samples_Metadata.csv"))



```

# Subset count table 

```{r}

# Select samples for MP Biofilm, filter out Mitochondria, chloroplast, Eukaryotas

count_table_subset_mx <- count_table %>% 
  select(MP_230809_B1_S1: MP_240616_B6_S54, Kingdom:Species) %>% 
  filter(Family != "Mitochondria" & Order != "Chloroplast" & Kingdom != "Eukaryota") %>% 
  filter(rowSums(across(where(is.numeric)))!=0) 
  


count_table_subset <- as.data.frame(count_table_subset_mx)







count_table_subset$c_o_f <- do.call(paste, c(count_table_subset[c("Class", "Order", "Family")], sep="_"))

count_table_subset$k_p <- do.call(paste, c(count_table_subset[c("Kingdom", "Phylum")], sep="_"))

count_table_subset$k_p_c <- do.call(paste, c(count_table_subset[c("Kingdom", "Phylum", "Class")], sep="_"))

count_table_subset$k_p_c_o_f_g_s <- do.call(paste, c(count_table_subset[c("Kingdom", "Phylum", "Class","Order", "Family","Genus", "Species")], sep="_"))

count_table_subset$taxa.ids <- ifelse(count_table_subset$c_o_f == "NA_NA_NA", count_table_subset$k_p, count_table_subset$c_o_f)

count_table_subset_ids <- count_table_subset %>% 
  select(c(Kingdom:taxa.ids)) 

count_table_subset_8000 <- count_table_subset %>% 
  select(-c(Kingdom:taxa.ids)) %>% 
   .[,colSums(.) >8000]
  

count_table_subset <- cbind(count_table_subset_8000, count_table_subset_ids)

count_table_subset_nosingle_doubles <- count_table_subset  %>% 
  filter(rowSums(across(where(is.numeric)))>2) 

# Singletons removes = 4355 - 3391 =  

```


# Aggregate by Family 


```{r}


agg_by_fam <- aggregate(count_table_subset_nosingle_doubles[,1:54], 
                        by = list(count_table_subset_nosingle_doubles$taxa.ids), 
                        FUN = sum)



```


# Aggregate by class

```{r}


agg_by_class <- aggregate(count_table_subset_nosingle_doubles[,1:54], 
                        by = list(count_table_subset_nosingle_doubles$k_p_c), 
                        FUN = sum)


```


# Agregate by Phylumn


```{r}


agg_by_phylum <- aggregate(count_table_subset_nosingle_doubles[,1:54], 
                        by = list(count_table_subset_nosingle_doubles$k_p), 
                        FUN = sum)



```

# Agregate by ASV


```{r}


agg_by_ASV <- aggregate(count_table_subset_nosingle_doubles[,1:54], 
                        by = list(count_table_subset_nosingle_doubles$k_p_c_o_f_g_s), 
                        FUN = sum)




```





# Calculate relative abundances by ASV

```{r}

# Count table with ASV's removing chloroplast, mitocondria and eukaryotes

count_table_subset_asv <- count_table_subset_nosingle_doubles %>% 
  filter(Family != "Mitochondria" & Order != "Chloroplast" & Kingdom != "Eukaryota") %>% 
  select(-c(Kingdom, Phylum, Class, Order, Family, Genus, Species, c_o_f, k_p, k_p_c, taxa.ids))

# Save Count table w/ ASV name as a csv

write_csv(count_table_subset_asv, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/count_table_subset_asv_forDEseq.csv")


# Calculate relative abundances 

numbers <- 1:5000

count_table_asv_RA <- as.data.frame(lapply(count_table_subset_asv[1:54], function(x) {(x/ sum(x))}))

count_table_asv_RA_asv <- cbind(count_table_subset_asv[55], count_table_asv_RA) #%>% 
  column_to_rownames(var = "k_p_c_o_f_g_s")

# Give ASV's unique identifiers and remove ASV's with no hits
  
count_table_asv_RA_asv_unique  <- count_table_asv_RA_asv %>%
  group_by(k_p_c_o_f_g_s) %>%
  mutate(k_p_c_o_f_g_s_unique = if (n() > 1) paste(k_p_c_o_f_g_s, numbers[row_number()], sep = "_") else as.character(k_p_c_o_f_g_s)) %>%
  ungroup() %>% 
   select(-k_p_c_o_f_g_s) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "k_p_c_o_f_g_s_unique") 

count_table_asv_RA_asv_unique_long <- count_table_asv_RA_asv_unique %>% 
  t()



write_csv(as.data.frame(count_table_asv_RA_asv_unique_long),"/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/count_table_asv_RA_asv_unique_long.csv")



```


# Calculate RA by Family 


```{r}


# 1. Calculate relative abundance:



family_RA <-  as.data.frame(lapply(agg_by_fam[2:length(agg_by_fam)], function(x) {(x/ sum(x))}))

# 2. Combine taxa IDs and relative abundance:

family_RA_w_taxa <- cbind(agg_by_fam[1], family_RA) %>% 
  remove_rownames %>%
  column_to_rownames(var = "Group.1") %>% 
   filter(rowSums(across(where(is.numeric)))!=0)


  

# 3. Transpose:

family_RA_taxa_tr <- data.frame(t(family_RA_w_taxa))






write_csv(family_RA_taxa_tr, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/family_RA_taxa_tr.csv")



family_RA_taxa_tr_for_sample_merger <- family_RA_taxa_tr %>% 
  rownames_to_column(var = "DNA_SampleID")

family_RA_w_metadata <- merge(sampletable_df_nmds, family_RA_taxa_tr_for_sample_merger, by = "DNA_SampleID")

family_RA_w_metadata <- family_RA_w_metadata %>% 
    mutate(phys_state = case_when(
    age_d < 50 ~ "Mature",
    age_d >= 50 ~ "Senescent"
  ))

family_RA_w_metadata %>% 
  filter(age_d == 16) %>% 
  group_by(phys_state, season) %>% 
  summarise(mean_Pirellula = mean(Planctomycetes_Pirellulales_Pirellulaceae),
            sd_Pirellula = sd(Planctomycetes_Pirellulales_Pirellulaceae))

abundant_family_RA 



```

# Stats: Planctomycetes_Pirellulales_Pirellulaceae RA between Spring and Summer at 16 days of age

```{r}

pirellula_16d <- family_RA_w_metadata %>% 
  filter(age_d == 16)

t.test(pirellula_16d$Planctomycetes_Pirellulales_Pirellulaceae[pirellula_16d$season == "Summer"],
            pirellula_16d$Planctomycetes_Pirellulales_Pirellulaceae[pirellula_16d$season == "Spring"],
            alternative = "less") #p = 0.03728, significant 




```




# Plot Major Families RA over time : Boxplots 


```{r}


pirellulales_16d_family <- family_RA_w_metadata %>% 
  ggplot(aes(x = age_d, y = Planctomycetes_Pirellulales_Pirellulaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Pirellulaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


granuloss_family_16days <- family_RA_w_metadata %>% 
  filter(age_d == 16) %>% 
  ggplot(aes(x = season, y = Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Granulosicoccaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 



flavobacteria_16d_fmaily <- family_RA_w_metadata %>% 
  filter(age_d == 16) %>% 
  ggplot(aes(x = season, y = Bacteroidia_Flavobacteriales_Flavobacteriaceae*100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Flavobacteriaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


```



```{r}


family_RA_taxa_tr_for_sample_merger <- family_RA_taxa_tr %>% 
  rownames_to_column(var = "DNA_SampleID")

family_RA_w_metadata <- merge(sampletable_df_nmds, family_RA_taxa_tr_for_sample_merger, by = "DNA_SampleID")



Pirellulaceae_family_p <- family_RA_w_metadata %>% 
  group_by(age_d, season) %>% 
  summarise(mean_RA = mean(Planctomycetes_Pirellulales_Pirellulaceae)*100,
            sd_RA = sd(Planctomycetes_Pirellulales_Pirellulaceae)*100) %>% 
  ggplot(aes(x = age_d, y = mean_RA, color = season)) +
  geom_point(size = 7) + 
  geom_line(aes(color = season)) +
  geom_errorbar(aes(ymin = mean_RA - sd_RA, ymax = mean_RA + sd_RA, width = 0)) +
  labs(x = "Age (d)",
       y = "Family Pirellulaceae (RA %)",
       color = "") +
  scale_y_continuous(breaks = c(0,10,20,30), limits = c(0,30)) +
  scale_color_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


Granulosicoccaceae_family_p <- family_RA_w_metadata %>% 
  group_by(age_d, season) %>% 
  summarise(mean_RA = mean(Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae)*100,
            sd_RA = sd(Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae)*100) %>% 
  ggplot(aes(x = age_d, y = mean_RA, color = season)) +
  geom_point(size = 7) + 
  geom_line(aes(color = season)) +
  geom_errorbar(aes(ymin = mean_RA - sd_RA, ymax = mean_RA + sd_RA, width = 0)) +
  labs(x = "Age (d)",
       y = "Family Granulosicoccaceae (RA %)",
       color = "") +
  scale_y_continuous(breaks = c(0,25,50,75,100), limits = c(0,75)) +
  scale_color_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


Thiotrichaceae_family_boxplot <- family_RA_w_metadata %>% 
  mutate(age_d = as.factor(age_d)) %>% 
  ggplot(aes(x = age_d, y = Gammaproteobacteria_Thiotrichales_Thiotrichaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Thiotrichaceae (RA %)",
       fill = "") +
  scale_color_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 

Flavobacteriaceae_family_p <- family_RA_w_metadata %>% 
  group_by(age_d, season) %>% 
  summarise(mean_RA = mean(Bacteroidia_Flavobacteriales_Flavobacteriaceae)*100,
            sd_RA = sd(Bacteroidia_Flavobacteriales_Flavobacteriaceae)*100) %>% 
  ggplot(aes(x = age_d, y = mean_RA, color = season)) +
  geom_point(size = 7) + 
  geom_line(aes(color = season)) +
  geom_errorbar(aes(ymin = mean_RA - sd_RA, ymax = mean_RA + sd_RA, width = 0)) +
  labs(x = "Age (d)",
       y = "Family Flavobacteriaceae (RA %)",
       color = "") +
  scale_y_continuous(breaks = c(0,5,10,15,20), limits = c(0,20)) +
  scale_color_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


Flavobacteriaceae_family_boxplot <- family_RA_w_metadata %>% 
  mutate(age_d = as.factor(age_d)) %>% 
  ggplot(aes(x = age_d, y = Bacteroidia_Flavobacteriales_Flavobacteriaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Flavobacteriaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 


Microtrichaceae_family_boxplot <- family_RA_w_metadata %>% 
  mutate(age_d = as.factor(age_d)) %>% 
  ggplot(aes(x = age_d, y = Acidimicrobiia_Microtrichales_Microtrichaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Microtrichaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 



Trueperaceae_family_boxplot <- family_RA_w_metadata %>% 
  mutate(age_d = as.factor(age_d)) %>% 
  ggplot(aes(x = age_d, y = Deinococci_Deinococcales_Trueperaceae *100, fill = season)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Family Trueperaceae (RA %)",
       fill = "") +
  scale_fill_manual(values = c("red", "black")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 











```

# Planctomycetes Pirellulales Family Mature vs. Senescent 


```{r}

pirellulales_family_sum <- family_RA_w_metadata %>% 
  select(age_d, season, Planctomycetes_Pirellulales_Pirellulaceae) %>% 
  mutate(phys_state = case_when(
    age_d < 50 ~ "Mature",
    age_d > 50 ~ "Senescent"
  )) 

boxplot_pirellulalaceae_phys_state <-  pirellulales_family_sum %>% 
  ggplot() +
  geom_boxplot(aes(x = phys_state, y = Planctomycetes_Pirellulales_Pirellulaceae*100, fill = phys_state)) +
  geom_jitter(aes(x = phys_state, y = Planctomycetes_Pirellulales_Pirellulaceae*100), width = 0.1) +
  labs(x = "",
       y = "Family Pirellulaceae (RA %)",
       fill = "") +
  scale_y_continuous(breaks = c(0,10,20,30), limits = c(0,30)) +
  scale_fill_manual(values = c("#034A85", "#9D3709"  )) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = "none",
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 



# Two Sample t-test: C:N content of mature kelp in Spring vs. Summer


## Test for Normality


with(pirellulales_family_sum, shapiro.test(Planctomycetes_Pirellulales_Pirellulaceae[phys_state== "Mature"])) # p = 0.77 = normal


with(pirellulales_family_sum, shapiro.test(Planctomycetes_Pirellulales_Pirellulaceae[phys_state== "Senescent"])) # p = 0.2042 = normal

## Equal Variances 

pirellulaes.ftest <- var.test(Planctomycetes_Pirellulales_Pirellulaceae ~ phys_state, data = pirellulales_family_sum)

pirellulaes.ftest  # Variances are equal. Use two sample t.test




pirellulales_t.test <- t.test(Planctomycetes_Pirellulales_Pirellulaceae ~ phys_state, data = pirellulales_family_sum, alternative = "greater")


pirellulales_t.test # p = 4.455e-05, df = 51.056, t = 4.2559

# Mean mature RA = 17.3%
# Mean senescent RA = 10.1%
 
# Relative abundance of Pirellulales family on mature kelp blades is significantly greater for mature blades than senescent blades



```

# Subset to 10 Families and Compare RA between Mature and Senescent Populations 


```{r}


family_RA_phys_long_top_families <- family_RA_w_metadata %>% 
  select(-c(DNA_SampleID:season, Chl_C)) %>% 
  melt(id = c("blade", "age_d")) %>% 
  arrange(desc(value)) %>% 
  filter(value > 0.01) %>% 
  mutate(phys_state = case_when(
    age_d < 50 ~ "Mature",
    age_d >= 50 ~ "Senescent"
  )) 



family_RA_phys_long_top_families_top10 <- family_RA_phys_long_top_families %>% 
  filter(variable %in% c("Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae", "Planctomycetes_Pirellulales_Pirellulaceae", "Gammaproteobacteria_Thiotrichales_Thiotrichaceae", "Bacteroidia_Flavobacteriales_Flavobacteriaceae", "Bacteroidia_Chitinophagales_Saprospiraceae", "Acidimicrobiia_Microtrichales_Microtrichaceae", "Alphaproteobacteria_Caulobacterales_Hyphomonadaceae", "Gammaproteobacteria_Arenicellales_Arenicellaceae", "Alphaproteobacteria_Rhodobacterales_Rhodobacteraceae", "Gammaproteobacteria_Pseudomonadales_Cellvibrionaceae")) %>% 
  separate(variable, "_", into = c("phylum", "order","family")) %>% 
  mutate(RA = as.numeric(value)) %>%
  select(-value)




top10_families_plot <- family_RA_phys_long_top_families_top10 %>% 
  ggplot() +
  geom_boxplot(aes(y = reorder(family, RA), x = RA*100, fill = phys_state)) +
  scale_fill_manual(values = c("#266A25","#BD9C3D")) +
  labs(y = "",
       x = "Relative Abundance (%)",
       fill = "") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.position = "top",
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 24),
        text = element_text(family = "Times")) 



```

# Subset to 10 Families and Compare RA between Seasons

```{r}


family_RA_phys_long_top_families_season <- family_RA_w_metadata %>% 
  select(-c(DNA_SampleID:sample_date, Chl_C, phys_state)) %>% 
  melt(id = c("blade", "season")) %>% 
  arrange(desc(value)) %>% 
  filter(value > 0.01)



family_RA_phys_long_top_families_top10_season <- family_RA_phys_long_top_families_season %>% 
  filter(variable %in% c("Gammaproteobacteria_Granulosicoccales_Granulosicoccaceae", "Planctomycetes_Pirellulales_Pirellulaceae", "Gammaproteobacteria_Thiotrichales_Thiotrichaceae", "Bacteroidia_Flavobacteriales_Flavobacteriaceae", "Bacteroidia_Chitinophagales_Saprospiraceae", "Acidimicrobiia_Microtrichales_Microtrichaceae", "Alphaproteobacteria_Caulobacterales_Hyphomonadaceae", "Gammaproteobacteria_Arenicellales_Arenicellaceae", "Alphaproteobacteria_Rhodobacterales_Rhodobacteraceae", "Gammaproteobacteria_Pseudomonadales_Cellvibrionaceae")) %>% 
  separate(variable, "_", into = c("phylum", "order","family")) %>% 
  mutate(RA = as.numeric(value)) %>%
  select(-value)




top10_families_plot_season <- family_RA_phys_long_top_families_top10_season %>% 
  ggplot() +
  geom_boxplot(aes(y = reorder(family, RA), x = RA*100, fill = season)) +
  scale_fill_manual(values = c("blue","red")) +
  labs(y = "",
       x = "Relative Abundance (%)",
       fill = "") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.position = "top",
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 24),
        text = element_text(family = "Times")) 



```

# Calculate RA by Class 



```{r}


# 1. Calculate relative abundance:

class_RA <-  as.data.frame(lapply(agg_by_class[2:length(agg_by_class)], function(x) {(x/ sum(x))}))

# 2. Combine taxa IDs and relative abundance:

class_RA_w_taxa <- cbind(agg_by_class[1], class_RA) %>% 
  remove_rownames %>%
  column_to_rownames(var = "Group.1")

# 3. Transpose:

class_RA_taxa_tr <- data.frame(t(class_RA_w_taxa))



write_csv(class_RA_taxa_tr, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/class_RA_taxa_tr.csv")





```


# Calculate RA by Phylum 



```{r}


# 1. Calculate relative abundance:

phylum_RA <-  as.data.frame(lapply(agg_by_phylum[2:length(agg_by_phylum)], function(x) {(x/ sum(x))}))

# 2. Combine taxa IDs and relative abundance:

phylum_RA_w_taxa <- cbind(agg_by_phylum[1], phylum_RA) %>% 
  remove_rownames %>%
  column_to_rownames(var = "Group.1")

# 3. Transpose:

phylum_RA_w_taxa_tr <- data.frame(t(phylum_RA_w_taxa))



write_csv(phylum_RA_w_taxa_tr, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/phylum_RA_taxa_tr.csv")


phylum_RA_taxa_tr_for_sample_merger <- phylum_RA_w_taxa_tr %>% 
  rownames_to_column(var = "DNA_SampleID")

phylum_RA_w_metadata <- merge(sampletable_df_nmds, phylum_RA_taxa_tr_for_sample_merger, by = "DNA_SampleID")

phylum_RA_w_metadata <- phylum_RA_w_metadata %>% 
    mutate(phys_state = case_when(
    age_d < 50 ~ "Mature",
    age_d >= 50 ~ "Senescent"
  ))


Phylum_summary_dominant <- phylum_RA_w_metadata %>% 
  summarize(mean_Bacteroides = mean(Bacteria_Bacteroidota),
            sd_Bacteroides = sd(Bacteria_Bacteroidota),
            mean_Proteobacteira = mean(Bacteria_Proteobacteria),
            sd_Proteobacteria = sd(Bacteria_Proteobacteria),
            mean_Planctomycetes = mean(Bacteria_Planctomycetota),
            sd_Planctomycetes = sd(Bacteria_Planctomycetota),
            mean_Verrucomicrobia = mean(Bacteria_Verrucomicrobiota),
            sd_Verrucomicrobia = sd(Bacteria_Verrucomicrobiota))

Phylum_summary_phys <- phylum_RA_w_metadata %>% 
  group_by(phys_state) %>% 
  summarize(mean_Bacteroides = mean(Bacteria_Bacteroidota),
            sd_Bacteroides = sd(Bacteria_Bacteroidota),
            mean_Proteobacteira = mean(Bacteria_Proteobacteria),
            sd_Proteobacteria = sd(Bacteria_Proteobacteria),
            mean_Planctomycetes = mean(Bacteria_Planctomycetota),
            sd_Planctomycetes = sd(Bacteria_Planctomycetota),
            mean_Verrucomicrobia = mean(Bacteria_Verrucomicrobiota),
            sd_Verrucomicrobia = sd(Bacteria_Verrucomicrobiota))

Phylum_summary_season <- phylum_RA_w_metadata %>% 
  group_by(season) %>% 
  summarize(mean_Bacteroides = mean(Bacteria_Bacteroidota),
            sd_Bacteroides = sd(Bacteria_Bacteroidota),
            mean_Proteobacteira = mean(Bacteria_Proteobacteria),
            sd_Proteobacteria = sd(Bacteria_Proteobacteria),
            mean_Planctomycetes = mean(Bacteria_Planctomycetota),
            sd_Planctomycetes = sd(Bacteria_Planctomycetota),
            mean_Verrucomicrobia = mean(Bacteria_Verrucomicrobiota),
            sd_Verrucomicrobia = sd(Bacteria_Verrucomicrobiota))

```

# Transform RA data: Arcsine transformation ASV Level


```{r}



ASV_RA_asv_transformed <- asin(sqrt(count_table_asv_RA_asv_unique))

# colnames(ASV_RA_asv_transformed) <- paste("asinsqrt", colnames(ASV_RA_asv_transformed), sep = "_")

ASV_RA_asv_transformed_tposed <- data.frame(t(ASV_RA_asv_transformed))

write_csv(ASV_RA_asv_transformed_tposed, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/ASV_RA_asv_transformed_tposed.csv")

ASV_RA_asv_transformed_tposed <- read_csv("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/ASV_RA_asv_transformed_tposed.csv")

```



# Transform RA data: Arcsine transformation Family Level



```{r}



family_RA_asv_transformed <- asin(sqrt(family_RA))

family_RA_asv_transformed_taxa <- cbind(agg_by_fam[1], family_RA_asv_transformed) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "Group.1")

family_RA_asv_transformed_taxa_tposed <- data.frame(t(family_RA_asv_transformed_taxa))

write_csv(family_RA_asv_transformed_taxa_tposed, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/family_RA_asv_transformed_taxa_tposed.csv")

```


# Transform RA data: Arcsine transformation Class Level



```{r}

class_RA_asv_transformed <- asin(sqrt(class_RA))

class_RA_asv_transformed_taxa <- cbind(agg_by_class[1], class_RA_asv_transformed) %>% 
  remove_rownames %>% 
  column_to_rownames(var = "Group.1")


class_RA_asv_transformed_taxa_tposed <- data.frame(t(class_RA_asv_transformed_taxa))

write_csv(class_RA_asv_transformed_taxa_tposed, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/class_RA_asv_transformed_taxa_tposed.csv")

```


# NMDS : ASV's: No singeltons or doublets 



```{r}

# Create a matrix of the Arcsine square root transformed RA 

com <- ASV_RA_asv_transformed_tposed[,1:ncol(ASV_RA_asv_transformed_tposed)]

m_com <- as.matrix(com)

# Calculate NMDS scores based on bray curtis dissimilarity 

nmds = metaMDS(m_com, dsitance = "bray")

data.scores = as.data.frame(scores(nmds)) %>% 
  rownames_to_column(var = "DNA_SampleID")


```


# Merge NMDS Scores with Sample Metadata


```{r}


sampletable_df_nmds <- read_csv(here("Inputdata", "sampledata", "MP_Biofilm_Samples_Metadata.csv"))


MP_Biofilm_Sample_Data_NMDS <- merge(sampletable_df_nmds, data.scores, by = "DNA_SampleID")

# Save NMDS and metadata merged file as a csv

write_csv(MP_Biofilm_Sample_Data_NMDS, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/MP_Biofilm_Sample_Data_NMDS_scores.csv")

# Load NMDS metadata merged file as a csv 

MP_Biofilm_Sample_Data_NMDS <- read_csv("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/MP_Biofilm_Sample_Data_NMDS_scores.csv")

# Plot results of community composition across potential density 

NMDS_p_biofilm <- MP_Biofilm_Sample_Data_NMDS %>% 
  ggplot(aes(x = NMDS1, y = NMDS2, color = phys_state, shape = season)) +
  geom_point(size = 7) +
  scale_color_manual(values = c("#266A25","#BD9C3D")) +
  labs(color = "",
       shape = "") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.position = "none",
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 



# scico::scale_color_scico(palette = "vik", breaks = c(1, 15,30,45,60,75, 90)) +

```


# Stats: PERMANOVA on Betadiversity between seasons and physiological state


```{r}

sampletable_df <- MP_Biofilm_Sample_Data_NMDS %>% 
  mutate(phys_state = case_when(
    age_d <50 ~ "Mature",
    age_d >=50 ~ "Senescent"
  )) %>% 
  select(DNA_SampleID,
         blade,
         season,
         age_d,
         phys_state,
         tissue_CN
         )


# PERMANOCVA community composition. Is there a significant difference in ASV relative abundance between mature and senescent kelp controlling for season (strata)? 

MP_Biofilm_ASB_adonis_phys_state <- adonis(ASV_RA_asv_transformed_tposed~phys_state, data = sampletable_df, method = "bray", strata = sampletable_df$season)


MP_Biofilm_ASB_adonis_phys_state # R2 = 0.18 p < 0.001 ***



# PERMANOCVA community composition. Is there a significant difference in ASV relative abundance between seasons controlling for phys_state (strata)? 

MP_Biofilm_ASB_adonis_season <- adonis(ASV_RA_asv_transformed_tposed~season, data = sampletable_df, method = "bray", strata = sampletable_df$phys_state)

MP_Biofilm_ASB_adonis_season # R2 = 0.26 P < 0.001 ***


# PERMANOCVA community composition. Is there a significant difference in ASV relative abundance between seasons. Just mature kelp. 

sampletable_df_mature <- sampletable_df %>% 
  filter(phys_state == "Mature")

mature_rows <- c(1:18, 31:42)

ASV_RA_asv_transformed_tposed_mature <- ASV_RA_asv_transformed_tposed[mature_rows, ]

MP_Biofilm_ASB_adonis_mature_season <- adonis(ASV_RA_asv_transformed_tposed_mature~season, data = sampletable_df_mature, method = "bray")

MP_Biofilm_ASB_adonis_mature_season


# PERMANOVA Season * Phys State 

# PERMANOCVA community composition. Is there a significant difference in ASV relative abundance between mature and senescent kelp controlling for season (strata)? 

MP_Biofilm_ASB_adonis_phys_state_seasons <- adonis(ASV_RA_asv_transformed_tposed~phys_state*season, data = sampletable_df, method = "bray")


MP_Biofilm_ASB_adonis_phys_state_seasons # R2 = 0.1 p < 0.001 ***


```


# Create Phyloseq Object 


```{r}

# Load counts table for MP Biofilm Samples and Remove ASV's that are not present 

count.table.dirty <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/seqtable_nochim.rds")

count.table.df <- as.data.frame(count.table.dirty)

count.table.df_subset <- count.table.df %>% 
  select(MP_230809_B1_S1:MP_240616_B6_S54) %>% 
  filter(rowSums(across(where(is.numeric)))>10) 
  
count.table <- as.matrix(count.table.df_subset)


taxa.table <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/taxa.rds")

sample.table <- readRDS("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sample.data.rds") %>% 
  drop_na(DNA_SampleID) %>% 
  column_to_rownames(var = "DNA_SampleID")


OTU = otu_table(count.table, taxa_are_rows = TRUE)
TAX = tax_table(taxa.table)
SAM = sample_data(sample.table)

ps = phyloseq(OTU, TAX, SAM)


ps_sub <- ps %>% 
  subset_taxa(Family != "Mitochondria", Class != "Chloroplast")



```


# DEseq2 : Physiological State 


```{r}

dd_phys_state <- phyloseq_to_deseq2(ps_sub, ~phys_state)

dd_phys_state$phys_state <- factor(dd_phys_state$phys_state, levels = c("Mature", "Senescent"))


dd_phys_state <- DESeq(dd_phys_state, test = "Wald", sfType = "poscounts")

resultsNames(dd_phys_state)


res = results(dd_phys_state, cooksCutoff = FALSE)
alpha = 0.01
sigtab = res[which(res$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(ps_sub)[rownames(sigtab), ], "matrix"))

head(sigtab)

dim(sigtab)


# Save Deseq2 Results for Dev Stage 

write_csv(sigtab, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/Deseq2_sigtab_dev_stage.csv")

# 

sigtab_dev <- sigtab %>% 
  rownames_to_column(var = "sequence")

# Phylum order
x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))

ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

# Pull out sequences of ASVs that have a significant difference between Phys States 

sigtab_phys_state_sequences <- sigtab %>% 
  rownames_to_column(var = "sequence")

sigtab_phys_state_sequence_list <- sigtab_phys_state_sequences$sequence




```


DEseq2: Season 


```{r}

dd_season <- phyloseq_to_deseq2(ps_sub, ~season)

dd_season$season <- factor(dd_season$season, levels = c("Summer", "Spring"))


dd_season <- DESeq(dd_season, test = "Wald", sfType = "poscounts")

resultsNames(dd_season)


res_season = results(dd_season, cooksCutoff = FALSE)

alpha = 0.01

sigtab_season = res_season[which(res_season$padj < alpha), ]

sigtab_season = cbind(as(sigtab_season, "data.frame"), as(tax_table(ps_sub)[rownames(sigtab_season), ], "matrix"))
head(sigtab)

dim(sigtab_season)

# Save Deseq2 Results for Seasons 

write_csv(sigtab_season, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/Deseq2_sigtab_season.csv")

#

sigtab_season_rownames <- sigtab_season %>% 
  rownames_to_column(var = "sequence")

x = tapply(sigtab_season$log2FoldChange, sigtab_season$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtab_season$Phylum = factor(as.character(sigtab_season$Phylum), levels=names(x))
# Genus order
x = tapply(sigtab_season$log2FoldChange, sigtab_season$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtab_season$Genus = factor(as.character(sigtab_season$Genus), levels=names(x))

ggplot(sigtab_season, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=6) + 
  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))


# Pull out sequences of ASVs that have a significant difference between Summer and Spring 

sigtab_season_sequences <- sigtab_season %>% 
  rownames_to_column(var = "sequence")

sigtab_season_sequence_list <- sigtab_season_sequences$sequence




```

# Merge Deseq2 Tables for Both

```{r}

sig_table_dev_season <- merge(sigtab_dev, sigtab_season_rownames, by = "sequence")


write_csv(sig_table_dev_season, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/Deseq2_sigtab_season_dev_combined.csv")

```


# Significant Sequences 

```{r}

sig_sequences <- unique(c(sigtab_phys_state_sequence_list, sigtab_season_sequence_list))

sig

```


# Filter ASVs to contain only those with signifcant differences between season and/or physiological state


```{r}

# Select Raw Counts from MP Biofilm Sampling and remove ASV singletons and doublets

count_table_MP_Biofilm <- count_table %>% 
  select(sequence, MP_230809_B1_S1:MP_240616_B6_S54, Kingdom:Species) %>% 
  filter(Family != "Mitochondria", Class != "Chloroplast") %>% 
  filter(rowSums(across(where(is.numeric)))>2) 

# Extract sequences and taxonomy, rename sequence column to Column

MP_Biofilm_sequences_taxa <- count_table_MP_Biofilm %>% 
  select(sequence, Kingdom:Species) %>% 
  mutate(Sequence = sequence) %>% 
  select(-sequence)



numeric_rows <- sapply(count_table_MP_Biofilm, is.numeric)

# Calculate RA's of Taxa

count_table_MP_Biofilm_RA <- as.data.frame(lapply(count_table_MP_Biofilm[2:55], function(x) {(x/ sum(x))}))

rownames(count_table_MP_Biofilm_RA) <- count_table_MP_Biofilm$sequence

# Filter out Amplicons that had a significant difference between mature, senescent kelp or between Summer and Spring 

count.table.df_subset_w_Sequences <- count_table_MP_Biofilm_RA %>% 
  rownames_to_column(var = "Sequence") %>% 
  filter(Sequence %in% c(sig_sequences)) 


# Merge Significant taxa RA dataframe with the Sequence and Taxonomy dataframes


sig_taxa_RA <- merge(count.table.df_subset_w_Sequences, MP_Biofilm_sequences_taxa, by = "Sequence")  



# Create Unique Taxa IDs

numbers_sig_taxa <- 1:500


sig_taxa_RA$f_g_s <- do.call(paste, c(sig_taxa_RA[c( "Family","Genus", "Species")], sep="_"))



# Give ASV's unique identifiers and remove ASV's with no hits
  
sig_taxa_RA_unique  <- sig_taxa_RA %>%
  group_by(f_g_s) %>%
  mutate(f_g_s_unique = if (n() > 1) paste(f_g_s, numbers_sig_taxa[row_number()], sep = "_") else as.character(f_g_s)) %>%
  ungroup() %>% 
   select(-c(f_g_s, Sequence, Kingdom:Species)) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "f_g_s_unique") %>% 
  t()

sig_taxa_RA_df <- as.data.frame(sig_taxa_RA_unique)

# Find ASVs with greater than 1% RA in at least 6 the  54 samples

sig_taxa_RA_df_majority <- sig_taxa_RA_df %>% 
  keep(~ (sum(. >= 0.01) / length(.)) > 0.111)




```




# Load in Sugar Mol% Data 


```{r}

sugars <- read_csv(here("Outputdata", "MP_DOM_sugar.csv"))



sugars_subset <- sugars %>% 
  select(DNA_SampleID, Fucose:`Man-URA`) %>% 
  column_to_rownames(var = "DNA_SampleID")




```


# Correlation Matrix 

## First Subset DNA Samples with Sugar Composition


```{r}

colors =  c("#034A85", "#9D3709")

sugar_rows <- c(1:36, 49:54)


sig_taxa_RA_df_majority_4sugars <- sig_taxa_RA_df_majority[sugar_rows,]



sugars_sigASV_cor <- cor(sig_taxa_RA_df_majority_4sugars, sugars_subset, method = "spearman")

sugars_sigASV_cor <- as.data.frame(sugars_sigASV_cor)

# Filter to only include ASVs with at least one Spearman's correlation coefficeint >0.5

sugars_sigASV_cor_R250 <- sugars_sigASV_cor %>% 
  filter_all(any_vars(abs(.) >= 0.5))



# Move rownames to column to save dataframe as a .csv 

sugars_sigASV_cor_R250_csv <- sugars_sigASV_cor_R250 %>% 
  rownames_to_column(var = "ASV_unique")


write_csv(sugars_sigASV_cor_R250_csv, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sugars_sigASV_cor_R250_csv.csv")


```


## Create dataframe for ASV Phylumn Characterization for heatmap 


```{r}

# Contains fgs grouping and phylumn name

sig_taxa_RA_phylum  <- sig_taxa_RA %>%
  group_by(f_g_s) %>%
  mutate(f_g_s_unique = if (n() > 1) paste(f_g_s, numbers_sig_taxa[row_number()], sep = "_") else as.character(f_g_s)) %>%
  ungroup() %>% 
  select(Phylum, f_g_s_unique)


f_g_s_df <- sugars_sigASV_cor_R250 %>% 
  rownames_to_column(var = "f_g_s_unique") %>% 
  select(f_g_s_unique)


sig_taxa_RA_phylum_majority <- sig_taxa_RA_phylum %>% 
  filter(f_g_s_unique %in% c(f_g_s_df$f_g_s_unique)) %>% 
  column_to_rownames(var = "f_g_s_unique")

sig_taxa_RA_phylum_majority_csv <- sig_taxa_RA_phylum_majority %>% 
  rownames_to_column(var = "ASV_unique")

# Save significant ASVs with Spearman's R2 >0.5 with at least one sugar as cvs

write_csv(sig_taxa_RA_phylum_majority_csv, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sig_taxa_RA_phylum_majority_csv.csv")


```


## Create heat map of significant ASV's vs. Mol% of Sugars released by giant kelp


```{r}

library(ggcorrplot)
library(pheatmap)

ann_colors = list(
    Phylum = c("cornflowerblue","orange","red2","forestgreen","mediumorchid1"))

sugars_ASV_corr_plot <- pheatmap(sugars_sigASV_cor_R250,
         annotation_row = sig_taxa_RA_phylum_majority,
         annotation_names_row = F, 
         fontsize_row = 15, 
         fontsize_col = 15,
         angle_col = 45,
         fontfamily = "Times",
         color = c("#001260", "#011A66", "#01226B" ,"#012A70", "#023275", "#023A7B", "#034280", "#034A85", "#05538A","#0A5C90", "#136697",
                  "#1E6F9D", "#2B79A4", "#3A84AB", "#4A8FB2", "#5A9ABA", "#6AA4C1", "#7BAEC8", "#8CB9CF", "#9CC3D5", "#ADCDDD", "#BDD6E3",
                  "#CDDFE7", "#DDE5E8", "#E7E7E5", "#EEE2DB", "#EEDACE", "#EBD0C1", "#E7C6B2", "#E3BCA5", "#DEB197", "#DAA78A", "#D69E7C",
                  "#D29470", "#CE8B64", "#C98157", "#C5784B", "#C26E3F", "#BE6533", "#B85C28", "#B1501C", "#A84411", "#9D3709", "#912D05",
                  "#872406", "#7D1C06", "#731405", "#6A0D06", "#620606", "#590007"))



```


# Create Plot of RA for major ASV that line up with the heatmap 


```{r}

sig_taxa_RA_df_majority_4sugars_long <- sig_taxa_RA_df_majority_4sugars %>% 
  remove_rownames() %>% 
  pivot_longer(
    cols = Rubinisphaeraceae_NA_NA_1:Granulosicoccaceae_Granulosicoccus_NA_14,
    names_to = "ASV",
    values_to = "RA") %>% 
  filter(ASV != "NA")



sig_ASV_RA_summary <- sig_taxa_RA_df_majority_4sugars_long %>% 
  filter(ASV != "NA") %>% 
  group_by(ASV) %>% 
  summarize(meanRA = mean(RA),
            sd_RA = sd(RA),
            minRA = min(RA),
            maxRA = max(RA))

# Significant ASVs written to CSV 

write_csv(sig_ASV_RA_summary, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/sig_ASV_RA_summary.csv")


sig_taxa_RA_df_majority_4sugars_long$ASV <- factor(sig_taxa_RA_df_majority_4sugars_long$ASV,
                                                   levels = 
                                                     c("Pirellulaceae_Blastopirellula_NA_12",
           "Pirellulaceae_Blastopirellula_NA_7",
           "Granulosicoccaceae_Granulosicoccus_NA_11",
           "Pirellulaceae_Blastopirellula_NA_11",
           "Granulosicoccaceae_Granulosicoccus_NA_6",
           "Saprospiraceae_Portibacter_NA_5",
           "Thiotrichaceae_Leucothrix_NA_8",
           "Microtrichaceae_Sva0996 marine group_NA_8",
          "Trueperaceae_Truepera_NA_1",
          "Microtrichaceae_Sva0996 marine group_NA_1",
          "Cellvibrionaceae_NA_NA_6",
          "Flavobacteriaceae_Ulvibacter_NA_2",
          "Granulosicoccaceae_Granulosicoccus_NA_8",
          "Arenicellaceae_Arenicella_NA_6",
          "Arenicellaceae_Arenicella_NA_3",
          "Flavobacteriaceae_Algitalea_NA",
          "Microtrichaceae_Sva0996 marine group_NA_7"))

# Levels for Only ASV with a significant correlation to a sugar

sig_ASV_R250_levels= c("Pirellulaceae_Blastopirellula_NA_12",
           "Pirellulaceae_Blastopirellula_NA_7",
           "Granulosicoccaceae_Granulosicoccus_NA_11",
           "Pirellulaceae_Blastopirellula_NA_11",
           "Granulosicoccaceae_Granulosicoccus_NA_6",
           "Saprospiraceae_Portibacter_NA_5",
           "Thiotrichaceae_Leucothrix_NA_8",
           "Microtrichaceae_Sva0996 marine group_NA_8",
          "Trueperaceae_Truepera_NA_1",
          "Microtrichaceae_Sva0996 marine group_NA_1",
          "Cellvibrionaceae_NA_NA_6",
          "Flavobacteriaceae_Ulvibacter_NA_2",
          "Granulosicoccaceae_Granulosicoccus_NA_8",
          "Arenicellaceae_Arenicella_NA_6",
          "Arenicellaceae_Arenicella_NA_3",
          "Flavobacteriaceae_Algitalea_NA",
          "Microtrichaceae_Sva0996 marine group_NA_7")


# Levels for all significant abundant ASVs

sig_ASV_levels = rev(c("Saprospiraceae_NA_NA_34",
         "Pirellulaceae_Rubripirellula_NA_3",
         "Pirellulaceae_Blastopirellula_NA_5",
         "Saprospiraceae_Portibacter_NA_5",
         "Granulosicoccaceae_Granulosicoccus_NA_10",
         "Saprospiraceae_NA_NA_22",
         "Granulosicoccaceae_Granulosicoccus_NA_14",
         "Pirellulaceae_Blastopirellula_NA_1",
         "Granulosicoccaceae_Granulosicoccus_NA_13",
         "Granulosicoccaceae_Granulosicoccus_NA_6",
         "Hyphomonadaceae_Litorimonas_NA_2",
         "Microtrichaceae_Sva0996 marine group_NA_2",
         "Thiotrichaceae_Cocleimonas_NA_7",
         "Pirellulaceae_Blastopirellula_NA_11",
         "Granulosicoccaceae_Granulosicoccus_NA_11",
         "Pirellulaceae_Blastopirellula_NA_7",
         "Pirellulaceae_Blastopirellula_NA_12",
         "Granulosicoccaceae_Granulosicoccus_NA_8",
         "Flavobacteriaceae_Ulvibacter_NA_2",
         "Cellvibrionaceae_NA_NA_6",
         "Microtrichaceae_Sva0996 marine group_NA_1",
         "Trueperaceae_Truepera_NA_1",
         "Microtrichaceae_Sva0996 marine group_NA_8",
         "Thiotrichaceae_Leucothrix_NA_8",
         "Microtrichaceae_Sva0996 marine group_NA_7",
         "Flavobacteriaceae_Algitalea_NA",
         "Arenicellaceae_Arenicella_NA_3",
         "Arenicellaceae_Arenicella_NA_6",
         "Arenicellaceae_Arenicella_NA_8",
         "Granulosicoccaceae_Granulosicoccus_NA_2",
         "Granulosicoccaceae_Granulosicoccus_NA_4",
         "Thiotrichaceae_Cocleimonas_NA_4",
         "Colwelliaceae_Thalassotalea_NA_2",
         "Rubinisphaeraceae_NA_NA_2",
         "KI89A clade_NA_NA",
         "Saprospiraceae_NA_NA_21",
         "Granulosicoccaceae_Granulosicoccus_NA_7",
         "Rubinisphaeraceae_NA_NA_3",
         "Microtrichaceae_Sva0996 marine group_NA_6",
         "Microtrichaceae_Sva0996 marine group_NA_3",
         "Saprospiraceae_NA_NA_32",
         "Pirellulaceae_Blastopirellula_NA_2",
         "Rubinisphaeraceae_NA_NA_1",
         "Hyphomonadaceae_Hellea_NA_1"
         ))




RA_sigtaxat_p <- sig_taxa_RA_df_majority_4sugars_long %>% 
  filter(ASV != "NA") %>% 
  ggplot(aes(x = RA*100, y = ASV)) + 
  geom_point() +
  labs(y = "",
       x = "Relative Abundance (%)") +
  scale_x_log10(breaks = c(0.01,0.1,1,10,60)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text.y = element_blank(),
        axis.text = element_text(size = 15, color = "black"),
        axis.title = element_text(size = 20),
        text = element_text(family = "Times")) 
  


combined_p <- plot_grid(sugars_ASV_corr_plot, RA_sigtaxat_p)

```


# Merge Sugar df and Sig ASV Dataset from Deseq2 for correlation analysis. Create a dataframe of Spearman correlation coefficients and p-values for ASVs with a significant relationship between a sugar  


```{r}
library(lattice)
library(latticeExtra)
library(Hmisc)

sugars_subset_2 <- sugars_subset %>% 
  rownames_to_column(var = "DNA_SampleID")

sig_taxa_RA_df_majority_4sugars_2 <- sig_taxa_RA_df_majority_4sugars %>% 
  rownames_to_column(var = "DNA_SampleID")

sugar_ASV <- merge(sugars_subset_2, sig_taxa_RA_df_majority_4sugars_2, by = "DNA_SampleID")

sugar_ASV <- sugar_ASV %>% 
  column_to_rownames(var = "DNA_SampleID") %>% 
  as.matrix()


rcorr_res <- rcorr(sugar_ASV, type = "spearman")

rcorr_res

rcorr_r2 <- as.data.frame(rcorr_res$r)

rcorr_r2_ASV_Sugar <- rcorr_r2[1:9, 10:53]

rcorr_p <- as.data.frame(rcorr_res$P)

rcorr_P_ASV_Sugar <- rcorr_p[1:9, 10:53]

rcorr_p_sigASV <- rcorr_P_ASV_Sugar %>% 
  select(Pirellulaceae_Blastopirellula_NA_12 ,
           Pirellulaceae_Blastopirellula_NA_7,
           Granulosicoccaceae_Granulosicoccus_NA_11,
           Pirellulaceae_Blastopirellula_NA_11,
           Granulosicoccaceae_Granulosicoccus_NA_6,
           Saprospiraceae_Portibacter_NA_5,
           Thiotrichaceae_Leucothrix_NA_8,
         `Microtrichaceae_Sva0996 marine group_NA_8`,
          Trueperaceae_Truepera_NA_1,
          `Microtrichaceae_Sva0996 marine group_NA_1`,
          Cellvibrionaceae_NA_NA_6,
          Flavobacteriaceae_Ulvibacter_NA_2,
          Granulosicoccaceae_Granulosicoccus_NA_8,
          Arenicellaceae_Arenicella_NA_6,
          Arenicellaceae_Arenicella_NA_3,
          Flavobacteriaceae_Algitalea_NA,
        `Microtrichaceae_Sva0996 marine group_NA_7`) %>% 
   mutate_if(is.numeric, round, digits = 5)

rcorr_r2_sigASV <- rcorr_r2_ASV_Sugar %>% 
  select(Pirellulaceae_Blastopirellula_NA_12 ,
           Pirellulaceae_Blastopirellula_NA_7,
           Granulosicoccaceae_Granulosicoccus_NA_11,
           Pirellulaceae_Blastopirellula_NA_11,
           Granulosicoccaceae_Granulosicoccus_NA_6,
           Saprospiraceae_Portibacter_NA_5,
           Thiotrichaceae_Leucothrix_NA_8,
         `Microtrichaceae_Sva0996 marine group_NA_8`,
          Trueperaceae_Truepera_NA_1,
          `Microtrichaceae_Sva0996 marine group_NA_1`,
          Cellvibrionaceae_NA_NA_6,
          Flavobacteriaceae_Ulvibacter_NA_2,
          Granulosicoccaceae_Granulosicoccus_NA_8,
          Arenicellaceae_Arenicella_NA_6,
          Arenicellaceae_Arenicella_NA_3,
          Flavobacteriaceae_Algitalea_NA,
        `Microtrichaceae_Sva0996 marine group_NA_7`) %>% 
 mutate_if(is.numeric, round, digits = 2)

combined_R2_p_value <- mapply(function(x, y){paste(x,y, sep = ",")},rcorr_r2_sigASV ,rcorr_p_sigASV )

combined_R2_p_value <- as.data.frame(combined_R2_p_value)

rownames(combined_R2_p_value) <- c(rownames(rcorr_P_ASV_Sugar))


write_csv(combined_R2_p_value, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/combined_R2_p_value.csv")



```




# Get Order of ASVs from Heatmap 

```{r}

sig_taxa_RA_df_majority_4sugars %>% 
  select(Saprospiraceae_NA_NA_34,
         Pirellulaceae_Rubripirellula_NA_3,
         Pirellulaceae_Blastopirellula_NA_5,
         Saprospiraceae_Portibacter_NA_5,
         Granulosicoccaceae_Granulosicoccus_NA_10,
         Saprospiraceae_NA_NA_22,
         Granulosicoccaceae_Granulosicoccus_NA_14,
         Pirellulaceae_Blastopirellula_NA_1,
         Granulosicoccaceae_Granulosicoccus_NA_13,
         Granulosicoccaceae_Granulosicoccus_NA_6,
         Hyphomonadaceae_Litorimonas_NA_2,
         `Microtrichaceae_Sva0996 marine group_NA_2`,
         Thiotrichaceae_Cocleimonas_NA_7,
         Pirellulaceae_Blastopirellula_NA_11,
         Granulosicoccaceae_Granulosicoccus_NA_11,
         Pirellulaceae_Blastopirellula_NA_7,
         Pirellulaceae_Blastopirellula_NA_12,
         Granulosicoccaceae_Granulosicoccus_NA_8,
         Flavobacteriaceae_Ulvibacter_NA_2,
         Cellvibrionaceae_NA_NA_6,
         `Microtrichaceae_Sva0996 marine group_NA_1`,
         Trueperaceae_Truepera_NA_1,
         `Microtrichaceae_Sva0996 marine group_NA_8`,
         Thiotrichaceae_Leucothrix_NA_8,
         `Microtrichaceae_Sva0996 marine group_NA_7`,
         Flavobacteriaceae_Algitalea_NA,
         Arenicellaceae_Arenicella_NA_3,
         Arenicellaceae_Arenicella_NA_6,
         Arenicellaceae_Arenicella_NA_8,
         Granulosicoccaceae_Granulosicoccus_NA_2,
         Granulosicoccaceae_Granulosicoccus_NA_4,
         Thiotrichaceae_Cocleimonas_NA_4,
         Colwelliaceae_Thalassotalea_NA_2,
         Rubinisphaeraceae_NA_NA_2,
         `KI89A clade_NA_NA`,
         Saprospiraceae_NA_NA_21,
         Granulosicoccaceae_Granulosicoccus_NA_7,
         Rubinisphaeraceae_NA_NA_3,
         `Microtrichaceae_Sva0996 marine group_NA_6`,
         `Microtrichaceae_Sva0996 marine group_NA_3`,
         Saprospiraceae_NA_NA_32,
         Pirellulaceae_Blastopirellula_NA_2,
         Rubinisphaeraceae_NA_NA_1,
         Hyphomonadaceae_Hellea_NA_1
         )





```

# Barplot of ASVs w/ significant sugar correlations 

```{r}

sig_taxa_RA_df_majority_4sugars_samplenames <- sig_taxa_RA_df_majority_4sugars %>% 
  rownames_to_column(var = "DNA_SampleID")


sig_taxa_RA_df_majority_4sugars_long_wSampleName <- sig_taxa_RA_df_majority_4sugars_samplenames %>%
  pivot_longer(
    cols = Rubinisphaeraceae_NA_NA_1:Granulosicoccaceae_Granulosicoccus_NA_14,
    names_to = "ASV",
    values_to = "RA") %>% 
  filter(ASV != "NA")

sig_taxa_RA_df_majority_4sugars_long_wSampleName$ASV <- factor(sig_taxa_RA_df_majority_4sugars_long_wSampleName$ASV,
                                                   levels = 
                                                     c("Pirellulaceae_Blastopirellula_NA_12",
           "Pirellulaceae_Blastopirellula_NA_7",
           "Granulosicoccaceae_Granulosicoccus_NA_11",
           "Pirellulaceae_Blastopirellula_NA_11",
           "Granulosicoccaceae_Granulosicoccus_NA_6",
           "Saprospiraceae_Portibacter_NA_5",
           "Thiotrichaceae_Leucothrix_NA_8",
           "Microtrichaceae_Sva0996 marine group_NA_8",
          "Trueperaceae_Truepera_NA_1",
          "Microtrichaceae_Sva0996 marine group_NA_1",
          "Cellvibrionaceae_NA_NA_6",
          "Flavobacteriaceae_Ulvibacter_NA_2",
          "Granulosicoccaceae_Granulosicoccus_NA_8",
          "Arenicellaceae_Arenicella_NA_6",
          "Arenicellaceae_Arenicella_NA_3",
          "Flavobacteriaceae_Algitalea_NA",
          "Microtrichaceae_Sva0996 marine group_NA_7"))

sig_taxa_RA_df_majority_4sugars_long_wSampleName <- sig_taxa_RA_df_majority_4sugars_long_wSampleName %>% 
  filter(ASV != "NA")
  
sig_taxa_cumulativeRA_plot <- sig_taxa_RA_df_majority_4sugars_long_wSampleName %>% 
  group_by(DNA_SampleID) %>% 
  summarize(Sum_RA = sum(RA)*100) %>% 
  ggplot(aes(x = DNA_SampleID, y = Sum_RA)) +
  geom_col()+
  labs(y = "Cumulative RA (%)",
       x = "Sample")+
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text.x = element_text(size = 10, angle = 90),
        axis.text = element_text(size = 15, color = "black"),
        axis.title = element_text(size = 20),
        text = element_text(family = "Times")) 



```


# Estimate Diversity 



```{r}


count_table_richness <- count_table %>% 
  filter(Family != "Mitochondria" & Order != "Chloroplast" & Kingdom != "Eukaryota") %>%
  select(sequence, MP_230809_B1_S1: MP_240616_B6_S54) %>% 
  filter(rowSums(across(where(is.numeric)))>0) %>% 
  column_to_rownames(var = "sequence")
  




# Generate Phyloseq Object

OTU = otu_table(count_table_richness, taxa_are_rows = TRUE)
TAX = tax_table(taxa_table_diversity)


ps = phyloseq(OTU)


# Calcualte Richness Statistics 

richness_df <- estimate_richness(ps, measures = c("Chao1", "Shannon")) %>%
  rownames_to_column(., var = "DNA_SampleID")


MP_biofilm_samples_richness <- merge(sampletable_df_nmds, richness_df, by = "DNA_SampleID")

# Save richness and metadata merged file as a csv

write_csv(MP_biofilm_samples_richness, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/MP_biofilm_samples_richness.csv")

# Load richness and metadata merged file as a csv 

MP_biofilm_samples_richness <- read_csv("/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_Microbiome/Outputdata/MP_biofilm_samples_richness.csv")

# Polynomial regression Shannon Divewrsity vs. age 

## Spring

MP_biofilm_samples_richness_spring <- MP_biofilm_samples_richness %>% 
  filter(season == "Spring")
  

spring_shannon_div_age_model <- lm(Shannon ~ poly(age_d, 2, raw = TRUE), 
                data = MP_biofilm_samples_richness_spring)

summary(spring_shannon_div_age_model)

new_x_spring  <- seq(min(MP_biofilm_samples_richness_spring$age_d),
                         max(MP_biofilm_samples_richness_spring$age_d), 1)
prediction_spring <- exp(predict(spring_shannon_div_age_model, newdata = list(age_d = new_x_spring)))
polyn_line_spring   <- data.frame(x = new_x_spring, y = prediction_spring)


## Summer

MP_biofilm_samples_richness_summer <- MP_biofilm_samples_richness %>% 
  filter(season == "Summer")


summer_shannon_div_age_model <- lm(Shannon ~ poly(age_d, 2, raw = TRUE), 
                data = MP_biofilm_samples_richness_summer)


summary(summer_shannon_div_age_model)

summer_shannon_div_age_model_linear <- lm(Shannon ~ age_d, 
                data = MP_biofilm_samples_richness_summer)

summary(summer_shannon_div_age_model_linear)

new_x_summer  <- seq(min(MP_biofilm_samples_richness_summer$age_d),
                         max(MP_biofilm_samples_richness_summer$age_d), 1)
prediction_summer <- exp(predict(summer_shannon_div_age_model, newdata = list(age_d = new_x_summer)))
polyn_line_summer   <- data.frame(x = new_x_summer, y = prediction_summer)





Shannon_div_plot <- MP_biofilm_samples_richness%>%
  ggplot(aes(x = age_d, y = Shannon, shape = season)) +
  geom_point(size = 5) +
  labs(x = "Age (d)",
       y = "Shannon Diversity",
       shape = "") +
   stat_smooth(method = "lm", formula = y ~ poly(x, 2), size = 2, se = FALSE, color = "black") +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.position = c(0.75,0.25),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 






Chao1_div_plot <- MP_biofilm_samples_richness%>%
  ggplot(aes(x = age_d, y = Chao1, color = season)) +
  geom_point(size = 5) +
  labs(x = "Age (d)",
       y = "Chao1",
       color = "") +
   stat_smooth(aes(color = season), method = "lm", se = FALSE) +
   scale_color_manual(values = c("red", "black")) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.position = c(0.8, 0.25),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 





```

# Stats: Diversity Metrics over time 

```{r}

library(rstatix)

# Repeated measures ANOVA of Shannon Diversity  by seaso  

Shannon_df <- MP_biofilm_samples_richness %>% 
  mutate(phase = case_when(
    age_d >= 50 ~ "Senescent",
    age_d < 50 ~ "Mature"
  )) %>% 
  select(season, phase, Shannon) 

#

Shannon_df_senescent <- Shannon_df %>% 
  filter(phase == "Senescent")

# Difference in diversity in senescent phase kelp. Kruskall walls test as data is not normally distributed 

wilcox.test(Shannon_df_senescent$Shannon[Shannon_df_senescent$season == "Summer"],
            Shannon_df_senescent$Shannon[Shannon_df_senescent$season == "Spring"], 
            alternative = "two.sided")

# No significant difference 

# Difference in diversity in senescent phase kelp. Kruskall walls test as data is normally distributed 

Shannon_df_mature <- Shannon_df %>% 
  filter(phase == "Mature")

t.test(Shannon_df_mature$Shannon[Shannon_df_mature$season == "Summer"],
            Shannon_df_mature$Shannon[Shannon_df_mature$season == "Spring"], 
            alternative = "two.sided") # 

```

