---
title: "MP_Exudate"
author: "Chance English"
date: "11/30/2023"
output: html_document
---


```{r setup, include=FALSE, echo=F, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE, message = F, echo = F, warning = F)
```

```{r, include=FALSE, echo=F, warning=F, message=F}



library(tidyverse)
library(lmodel2)
library(ggplot2)
library(readxl)
library(kableExtra)
library(scales)
library(factoextra)
library(mgcv)
library(janitor)
library(broom)
library(here)
library(rstatix)
library(ggpubr)
library(vegan)
library(ggbump)
library(RColorBrewer)
library(scico)
library(sicegar)

```

# Load Data 

```{r, include=FALSE, echo=F, warning=F, message=F}


MP_DOC_NPP_phys_env_sum <- read_csv(here("Outputdata", "MP_DOC_NPP_phys_env_sum.csv"))





```


# Total DOC vs NPP 


```{r}


doc_NPP <- MP_DOC_NPP_phys_env_sum %>%
  filter(flag != 2) %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(x = NPP_umolC_g_hr, y = DOC_umolC_g_hr)) +
  geom_point(aes(color = age_d), size = 7) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "Age (d)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 2) +
  geom_abline(intercept = 0.82, slope = 0.023, size = 2) +
  scico::scale_color_scico(palette = "vik", breaks = c(0,25,50,75,100)) +
  scale_x_continuous(breaks = c(0,100,200,300), limits = c(-30,300)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.8, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 

 

PER_NPP <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2) %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(lightlevel_timepoint != 1) %>% 
  filter(DOC_umolC_g_hr > 0 & NPP_umolC_g_hr > 10) %>% 
  ggplot(aes(x = NPP_umolC_g_hr, y = (DOC_umolC_g_hr/NPP_umolC_g_hr)*100)) +
  geom_point(aes(color = age_d, shape = season), size = 7) +
  labs(y = "Percent Extracellular Release",
       shape = "") +
  xlab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  scale_shape_manual(values = c(19,17)) +
  labs(color = "Age (d)") +
  scale_y_log10() +
  scico::scale_color_scico(palette = "vik", breaks = c(0,25,50,75,100)) +
  scale_x_log10() +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.4, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times")) 




```


# Mature (not senescent) DOC vs NPP 


```{r}

doc_NPP_mature <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  filter(age_d <= 50) %>% 
  ggplot(aes(x = NPP_umolC_g_hr, y = DOC_umolC_g_hr, color = PAR)) +
  geom_point(size = 7) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "PAR") + 
  scale_y_continuous(labels = label_number(accuracy = 1), limits = c(0,10), breaks = c(0,5,10)) +
  scale_x_continuous(limits = c(-50,300),breaks = c(0,100,200,300)) +
   scale_color_viridis_b(breaks = c(0,50,100,250,500,1000,1500), limits = c(0,1500), name = expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  geom_abline(slope = 0.023, intercept = 0.82, size = 2) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.7, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

# Log 10 transformed rates 

doc_NPP_mature_log10 <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(PAR > 10) %>% 
  filter(flag != 2) %>% 
  filter(age_d <= 50) %>% 
  filter(log10(DOC_umolC_g_hr) >= -2) %>% 
  ggplot(aes(x = log10(NPP_umolC_g_hr), y = log10(DOC_umolC_g_hr))) +
  geom_point(size = 7) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "Age (d)") + 
  geom_smooth(method = "lm") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))






```



# Summary Stats DOC Release Rates by Season, Physiological State

```{r}

DOC_rate_sum <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2 & season %in% c("Spring", "Summer")) %>% 
  mutate(phys_state = case_when(
    age_d < 50 ~ "Mature",
    age_d >= 50 ~ "Senescent"
  )) %>%
  group_by(season, phys_state) %>% 
  select(DOC = DOC_umolC_g_hr) 

DOC_rate_phys_state_sum <- 
  DOC_rate_sum %>% 
  group_by(phys_state) %>% 
  summarize(mean_DOC = mean(DOC),
            sd_DOC = sd(DOC),
            min_DOC = min(DOC), 
            max_DOC = max(DOC))


doc_season_phys_state_boxplot <- DOC_rate_sum %>% 
  ggplot() +
  geom_boxplot(aes(x = season, y = DOC, fill = phys_state)) + 
  scale_fill_manual(values = c("#266A25","#BD9C3D")) +
  scale_color_manual(values = c("darkgreen","darkgoldenrod")) +
  labs(x = "") + 
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = "top",
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))


# Two Sample t-test: C:N content of mature kelp in Spring vs. Summer


## Test for Normality

mature_DOC <- DOC_rate_sum %>% 
  filter(phys_state == "Mature") 

with(mature_DOC, shapiro.test(DOC[season == "Summer"])) # p = 0.0129 = not-normal

with(mature_DOC, shapiro.test(DOC[season == "Spring"])) # p = 0.013 = not-normal

# DOC Release Rates Mature Kelp 

res_DOC_mature_season <- wilcox.test(mature_DOC$DOC[mature_DOC$season == "Summer"],mature_DOC$DOC[mature_DOC$season == "Spring"])

res_DOC_mature_season # p = 0.6836, no significant difference in DOC release rates summer vs spring 


## Test for Normality: Senescent Kelp by season

senescent_DOC <- DOC_rate_sum %>% 
  filter(phys_state == "Senescent") 

with(senescent_DOC, shapiro.test(DOC[season == "Summer"])) # p <0.001= not-normal

with(senescent_DOC, shapiro.test(DOC[season == "Spring"])) # p < 0.001 = not-normal

# DOC Release Rates Mature Kelp 

res_DOC_senescent_season <- wilcox.test(senescent_DOC$DOC[senescent_DOC$season == "Summer"],senescent_DOC$DOC[senescent_DOC$season == "Spring"])

res_DOC_senescent_season # p = 0.6836, no significant difference in DOC release rates summer vs spring 


## Test for Normality: Senescent Kelp by season: Summer

summer_DOC <- DOC_rate_sum %>% 
  filter(season == "Summer") 

with(summer_DOC, shapiro.test(DOC[phys_state == "Mature"])) # p <0.001= not-normal

with(summer_DOC, shapiro.test(DOC[phys_state == "Senescent"])) # p < 0.001 = not-normal

# DOC Release Rates Mature Kelp 

res_Summer_DOC_mature_v_senescent <- wilcox.test(summer_DOC$DOC[summer_DOC$phys_state == "Mature"],summer_DOC$DOC[summer_DOC$phys_state == "Senescent"])

res_Summer_DOC_mature_v_senescent # W = 189.5, p = 8.712e-10 ***


## Test for Normality: Senescent Kelp by season: Spring

spring_DOC <- DOC_rate_sum %>% 
  filter(season == "Spring") 

with(spring_DOC, shapiro.test(DOC[phys_state == "Mature"])) # p <0.001= not-normal

with(spring_DOC, shapiro.test(DOC[phys_state == "Senescent"])) # p < 0.001 = not-normal

# DOC Release Rates Mature Kelp 

res_Spring_DOC_mature_v_senescent <- wilcox.test(spring_DOC$DOC[spring_DOC$phys_state == "Mature"],spring_DOC$DOC[spring_DOC$phys_state == "Senescent"])

res_Spring_DOC_mature_v_senescent # W = 275, p = 2.709 e-05 ***



```


# Nitrate and dry mass Normalized NPP at saturating irradiance over age


```{r}

NPP_age_season_Pmax <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, PAR >= 400, season %in% c("Summer", "Spring")) %>% 
  group_by(age_d, season) %>% 
  summarize(NPP = mean(NPP_umolC_g_hr),
            NPP_sd = sd(NPP_umolC_g_hr),
            nitrate = mean(N_N_uM),
            nitrate_sd = mean(N_N_uM),
            NPP_nitrate = NPP/nitrate,
            NPP_nitrate_sd = sqrt((NPP_sd^2/NPP)+ (nitrate_sd^2/nitrate))) %>%
  ungroup() %>% 
  ggplot(aes(x = age_d, y = NPP, color = season)) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = NPP - NPP_sd, ymax = NPP + NPP_sd, color = season), width = 0 ) +
  geom_smooth(method = "lm", se = FALSE,aes(color = season)) +
  ylab(expression(P[max]~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  scale_color_manual(values = c("red", "black")) +
  labs(x = "Age (d)",
       color = "") + 
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(1, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 1),
        legend.position = c(0.8,0.75),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))


# NO3 normalized rates


NPP_age_season_Pmax <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, PAR >= 400, season %in% c("Summer", "Spring")) %>% 
  group_by(age_d, season) %>% 
  summarize(NPP = mean(NPP_umolC_g_hr),
            NPP_sd = sd(NPP_umolC_g_hr),
            nitrate = mean(N_N_uM),
            nitrate_sd = mean(N_N_uM),
            NPP_nitrate = NPP/nitrate,
            NPP_nitrate_sd = sqrt((NPP_sd^2/NPP)+ (nitrate_sd^2/nitrate))) %>%
  ungroup() %>% 
  ggplot(aes(x = age_d, y = NPP, color = season)) +
  geom_point(size = 7) +
  geom_errorbar(aes(ymin = NPP - NPP_sd, ymax = NPP + NPP_sd, color = season), width = 0 ) +
  geom_smooth(method = "lm", se = FALSE,aes(color = season)) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"} ~ NO["3"]^{"-"}^{"-1"}))) +
  labs(x = "Age (d)") + 
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(1, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))



```


# NPP over Age with PAR 


```{r}

MP_NPP_age_par <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  ggplot(aes(y = NPP_umolC_g_hr, x = age_d)) +
  geom_point(aes(color = PAR, shape = season), size = 7) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "PAR",
       x = "Age (d)",
       shape = "") + 
  scale_x_continuous(limits = c(10,90), breaks = c(25,50,75)) +
  scale_y_continuous(limits = c(-50,300), breaks = c(0,100,200,300)) +
  geom_hline(yintercept = 0, color = "black", size= 1, linetype = "dashed") +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        legend.key.height = unit(0.7, "inch"),
        legend.key.width = unit(0.4, "inch"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times")) +
  scale_shape_manual(values = c(19,17)) +
  scale_color_viridis_b(breaks = c(0,50,100,250,500,1000,1500), limits = c(0,1500), name = expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  guides(shape = "none")





```

# OLS, Maximum Rates of NPP vs Age both seasons 

```{r}

# Spring 

MP_NPP_age_Spring <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  filter(lightlevel_timepoint %in% c(5,6), season == "Spring") %>% 
  select(age_d, NPP_umolC_g_hr)



OLS_NPP_age_Spring <- lm(NPP_umolC_g_hr ~ age_d, data = MP_NPP_age_Spring)
summary(OLS_NPP_age_Spring)


# Summer 

MP_NPP_age_Summer <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  filter(lightlevel_timepoint %in% c(5,6), season == "Summer") %>% 
  select(age_d, NPP_umolC_g_hr)

OLS_NPP_age_Summer <- lm(NPP_umolC_g_hr ~ age_d, data = MP_NPP_age_Summer)
summary(OLS_NPP_age_Summer)


# Plot Pmax vs Age for Both Seasons 

MP_Pmax_Age_p <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  filter(lightlevel_timepoint %in% c(5,6)) %>% 
  select(age_d, NPP_umolC_g_hr, season) %>% 
  ggplot(aes(y = NPP_umolC_g_hr, x = age_d)) +
  geom_point(aes(shape = season, color = season), size = 7) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "",
       x = "Age (d)",
       shape = "") +
  scale_color_manual(values = c("red", "black"), guide = "none") +
  geom_abline(slope = -3.22, intercept = 260.73, size = 2, color = "black") +
  geom_abline(slope = -1.34, intercept = 176.64, size = 2, color = "red") +
  scale_x_continuous(limits = c(10,80), breaks = c(0,25,50,75)) +
  scale_y_continuous(limits = c(-50,300), breaks = c(0,100,200,300)) +
  geom_hline(yintercept = 0, color = "black", size= 1, linetype = "dashed") +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        legend.key.height = unit(0.7, "inch"),
        legend.key.width = unit(0.4, "inch"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times")) +
  scale_shape_manual(values = c(19,17))


```



# NPP over Age with PAR 


```{r}

NPP_par <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(flag != 2) %>% 
  group_by(age_d, season, lightlevel_timepoint) %>% 
  mutate(mean_PAR = mean(PAR),
         mean_NPP = mean(NPP_umolC_g_hr),
         sd_PAR = sd(PAR),
         sd_NPP = sd(NPP_umolC_g_hr)) %>% 
  ggplot(aes(y = mean_NPP, x = mean_PAR, group = interaction(age_d, season))) +
  labs(color = "Age (d)") +
  geom_point(aes(shape = season, color = age_d), size = 7) +
  geom_line(aes(color = age_d), size = 1) +
  geom_errorbar(aes(xmin = mean_PAR - sd_PAR, xmax = mean_PAR + sd_PAR, ymin = mean_NPP - sd_NPP, ymax = mean_NPP + sd_NPP), width = 0, alpha = 0.1) +
  scale_shape_manual(values = c(19,17), guide = "none") +
  scico::scale_color_scico(palette = "vik", breaks = c(0,25,50,75,100)) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  scale_x_continuous(limits = c(0,1550), breaks = c(0,300,600,900,1200,1600)) +
  scale_y_continuous(limits = c(-50,300), breaks = c(0,100,200,300)) +
  geom_hline(yintercept = 0, color = "black", size= 1, linetype = "dashed") +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        legend.key.height = unit(0.7, "inch"),
        legend.key.width = unit(0.4, "inch"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times")) #
  scale_color_viridis_b(breaks = c(0,50,100,250,500,1000,1500), limits = c(0,1500)) #, name = expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  guides(shape = "none")



```

# Respiration Rates when PAR = 0

```{r}

respiration <- MP_DOC_NPP_phys_env_sum_flags %>% 
  filter(PAR == 0) %>% 
  select(age_d, season, NPP_umolC_g_hr)

respiration_spring <- respiration %>% 
  filter(season == "Spring")

respiration_summer <- respiration %>% 
  filter(season == "Summer")

spring_respiration_lm_age <- lm(NPP_umolC_g_hr ~ age_d, data = respiration_spring)

summary(spring_respiration_lm_age)

summer_respiration_lm_age <- lm(NPP_umolC_g_hr ~ age_d, data = respiration_summer)

summary(summer_respiration_lm_age)

```


# Summary Table for Each Sampling Period

```{r}

table_1_sum <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2 & season %in% c("Spring", "Summer")) %>% 
  group_by(season, date, age_d) %>% 
  summarize(n = 6, 
    blade = blade,
    lightlevel_timepoint = lightlevel_timepoint,
    mean_CN = round(mean(C_N),1),
    sd_CN = round(sd(C_N),1),
    min_PAR = round(min(PAR),0),
    max_PAR = round(max(PAR),0),
    mean_ChlC = round(mean(Chl_C*1000),1),
    sd_ChlC = round(sd(Chl_C *1000),1),
    min_DOC = round(min(DOC_umolC_g_hr),1),
    max_DOC = round(max(DOC_umolC_g_hr),1),
    min_NPP = round(min(NPP_umolC_g_hr),1),
    max_NPP = round(max(NPP_umolC_g_hr),1)
  ) %>% 
  filter(blade == 1 & lightlevel_timepoint == 1) %>% 
  select(-c(blade, lightlevel_timepoint)) %>% 
  mutate(CN = paste(mean_CN, sd_CN, sep = " ± "),
         ChlC = paste(mean_ChlC, sd_ChlC, sep = " ± "),
        PAR_range = paste(min_PAR, max_PAR, sep = " - "),
         NPP_range = paste(min_NPP, max_NPP, sep = " - "),
        DOC_range = paste(min_DOC, max_DOC, sep = " - ")) %>% 
  select(-c(min_DOC, max_DOC, 
            min_NPP, max_NPP, 
            min_PAR, max_PAR, 
            mean_CN, sd_CN, 
            mean_ChlC, sd_ChlC)) %>% 
  arrange(season, age_d)

table_1_sum


write_excel_csv(table_1_sum, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_DOC/Data/table_1_sum.csv")

table_1 <- kableExtra::kbl(table_1_sum,
                booktabs = TRUE,
                col.names = c("Season", 
                              "Date",
                              "Age (d)",
                              "N",
                              "Blade C:N",
                              "Blade Chl:C",
                              "$PAR~(µmol~m^{2}~s^{-1})$",
                              "$NPP~(µmolC~g_{DW}^{-1}~hr^{-1})$",
                              "$DOC~(µmolC~g_{DW}^{-1}~hr^{-1})$")) %>% 
  kable_classic(html_font = "Times New Roman",
                  full_width = F) %>% 
  kable_styling(font_size = 20)


```



# NLS exponential model fit for DOC production and age: Summer


```{r}

# Model DOC Spring and SUmmer over Age

MP_DOC_NPP_phys_env_SUMMER <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2) %>% 
  filter(lightlevel_timepoint %in% c(5,6)) %>% 
  filter(season == "Summer") %>% 
  group_by(age_d) %>% 
  summarize(DOC_p_mean = mean(DOC_umolC_g_hr),
            DOC_p_sd = sd(DOC_umolC_g_hr)) 



# Select an approximate $\theta$, since theta must be lower than min(y), and greater than zero
theta.0 <- min(MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean) *0.99 

# Estimate the rest parameters using a linear model
model.0 <- lm(log(DOC_p_mean - theta.0) ~ age_d, data=MP_DOC_NPP_phys_env_SUMMER)  
alpha.0 <- exp(coef(model.0)[1])
beta.0 <- coef(model.0)[2]

# Starting parameters
start <- list(alpha = alpha.0, beta = beta.0, theta = theta.0)
start

model_nls_SUMMER_DOC_age <- nls(DOC_p_mean ~ alpha * exp(beta * age_d) + theta , data = MP_DOC_NPP_phys_env_SUMMER, start = start)

plot(MP_DOC_NPP_phys_env_SUMMER$age_d, MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean)
lines(MP_DOC_NPP_phys_env_SUMMER$age_d, predict(model_nls_SUMMER_DOC_age, list(x= MP_DOC_NPP_phys_env_SUMMER$age_d)))





```


# NLS exponential model fit for DOC production and age: Spring


```{r}

# Model DOC Spring and SUmmer over Age

MP_DOC_NPP_phys_env_Spring <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2) %>% 
  filter(lightlevel_timepoint %in% c(5,6)) %>% 
  filter(season == "Spring") %>% 
  group_by(age_d) %>% 
  summarize(DOC_p_mean = mean(DOC_umolC_g_hr),
            DOC_p_sd = sd(DOC_umolC_g_hr)) 


# Select an approximate $\theta$, since theta must be lower than min(y), and greater than zero
theta.spring <- min(MP_DOC_NPP_phys_env_Spring$DOC_p_mean) *0.99 

# Estimate the rest parameters using a linear model
model.spring <- lm(log(DOC_p_mean - theta.spring) ~ age_d, data=MP_DOC_NPP_phys_env_SUMMER)  
alpha.spring <- exp(coef(model.spring)[1])
beta.spring <- coef(model.spring)[2]

# Starting parameters
start.spring <- list(alpha = alpha.spring, beta = beta.spring, theta = theta.spring)
start.spring

model_nls_Spring_DOC_age <- nls(DOC_p_mean ~ alpha * exp(beta * age_d) + theta , data = MP_DOC_NPP_phys_env_Spring, start = start.spring)

plot(MP_DOC_NPP_phys_env_Spring$age_d, MP_DOC_NPP_phys_env_Spring$DOC_p_mean) +
lines(MP_DOC_NPP_phys_env_Spring$age_d, predict(model_nls_Spring_DOC_age, list(x= MP_DOC_NPP_phys_env_Spring$age_d)))





```


# Plot DOC Production vs. Age for Spring and Summer w/ NLS Exponential Fits


```{r}


spring_summer_DOC_age_nls_fit_plot <- ggplot() +
  geom_point(aes(x = MP_DOC_NPP_phys_env_SUMMER$age_d, y = MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean,  color = "Summer"), size = 3) +
  geom_point(aes(x = MP_DOC_NPP_phys_env_Spring$age_d, y = MP_DOC_NPP_phys_env_Spring$DOC_p_mean, color = "Spring"), size = 3) +
  geom_smooth(aes(x = MP_DOC_NPP_phys_env_SUMMER$age_d, y = MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean),
    method = "nls", se = FALSE,
    formula = y ~ a * exp(b * x),
    method.args = list(start = c(a = 0.022, b = 0.0875)), color = "black") +
   geom_smooth(aes(x = MP_DOC_NPP_phys_env_Spring$age_d, y = MP_DOC_NPP_phys_env_Spring$DOC_p_mean),
    method = "nls", se = FALSE,
    formula = y ~ a * exp(b * x),
    method.args = list(start = c(a = 0.0235, b = 0.086)), color = "blue") +
  geom_errorbar(aes(x = MP_DOC_NPP_phys_env_Spring$age_d,
                    y = MP_DOC_NPP_phys_env_Spring$DOC_p_mean,
                    ymin = MP_DOC_NPP_phys_env_Spring$DOC_p_mean - MP_DOC_NPP_phys_env_Spring$DOC_p_sd, 
                     ymax =  MP_DOC_NPP_phys_env_Spring$DOC_p_mean + MP_DOC_NPP_phys_env_Spring$DOC_p_sd,
                    color = "Spring"),
                width = 0) +
  geom_errorbar(aes(x = MP_DOC_NPP_phys_env_SUMMER$age_d, 
                    y = MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean,
                    ymin = MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean - MP_DOC_NPP_phys_env_SUMMER$DOC_p_sd, 
                    ymax = MP_DOC_NPP_phys_env_SUMMER$DOC_p_mean + MP_DOC_NPP_phys_env_SUMMER$DOC_p_sd,
                    color = "Summer"), 
                width = 0) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "Season",
       x = "Age (d)") + 
  scale_color_manual(values = c("blue", "black"), labels = c("Spring", "Summer")) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))





```


# DOC Production vs Bulk C:N and Chl:C

```{r}

DOC_CN_Chl_C <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, season %in% c("Spring", "Summer")) %>% 
  select(age_d, season, DOC_umolC_g_hr, C_N, Chl_C, chlA_nmol_g)

DOC_CN_Chl_C %>% 
  ggplot(aes(x = C_N, y = DOC_umolC_g_hr)) +
  geom_point(aes(color = season))

DOC_tissueCN_lm <- lmodel2(DOC_umolC_g_hr ~ C_N, data = DOC_CN_Chl_C)

DOC_tissueCN_lm


  
```


# Mature DOC Production: Dark vs. Light Saturating

```{r}

Mature_DOC <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2,
         season %in% c("Spring", "Summer"),
         age_d < 50) %>% 
  select(PAR, DOC_umolC_g_hr) %>% 
  mutate(light_level = case_when(
    PAR == 0 ~ "Dark",
    PAR > 300 ~ "Saturating",
    PAR > 0 & PAR < 300 ~ "Limiting"
  ))


Mature_DOC %>% 
  ggplot(aes(x = light_level, y = DOC_umolC_g_hr)) +
  geom_boxplot() +
  theme_bw()
           
light_level_doc.aov <- aov(DOC_umolC_g_hr ~ light_level , data = Mature_DOC)

summary(light_level_doc.aov)

TukeyHSD(light_level_doc.aov)

plot(light_level_doc.aov, 1)

library(car)

leveneTest(DOC_umolC_g_hr ~ light_level, data = Mature_DOC) # no evidence to suggest that variances are signifciantly differnce between groups 


plot(light_level_doc.aov, 2)

aov_residuals <- residuals(object = light_level_doc.aov )

shapiro.test(x = aov_residuals) # Not normal, use Kruskal-Wallis

kruskal.test(DOC_umolC_g_hr ~ light_level, data = Mature_DOC)


```


# PER vs. CN

```{r}

PER_C_N <-  MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, season %in% c("Spring", "Summer")) %>% 
  select(age_d, season, PAR,DOC_umolC_g_hr,NPP_umolC_g_hr, C_N, Chl_C, chlA_nmol_g) %>% 
  filter(PAR > 0) %>% 
  filter(age_d < 50) %>% 
  mutate(PER = DOC_umolC_g_hr/NPP_umolC_g_hr *100) %>% 
  filter(NPP_umolC_g_hr > 0) %>% 
  ggplot(aes(x = C_N, y = PER)) +
  geom_point() +
  labs(y = "Percent Extracellular Release",
       x = "Tissue C:N") + 
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))


lm_PER_CN <- lmodel2(PER ~ C_N, data = PER_CN)

lm_PER_CN

PER_CN10 <- 10*-0.0468 + 3.4496
PER_CN40 <- 40*-0.0468 + 3.4496

```

# PER vs. PAR

```{r}

PER_PAR <-  MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, season %in% c("Spring", "Summer")) %>% 
  select(age_d, season, PAR,DOC_umolC_g_hr,NPP_umolC_g_hr, C_N, Chl_C, chlA_nmol_g) %>% 
  filter(PAR > 0) %>% 
  filter(age_d < 50) %>% 
  mutate(PER = DOC_umolC_g_hr/NPP_umolC_g_hr *100) %>% 
  filter(NPP_umolC_g_hr > 0) %>% 
  ggplot(aes(x = PAR, y = PER)) +
  geom_point() +
  labs(y = "Percent Extracellular Release",
       x = expression(paste("PAR (µmol ",m^-2, s^-1,")"))) + 
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

PER_PAR <-  MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2, season %in% c("Spring", "Summer")) %>% 
  select(age_d, season, PAR,DOC_umolC_g_hr,NPP_umolC_g_hr, C_N, Chl_C, chlA_nmol_g) %>% 
  filter(PAR > 0) %>% 
  filter(age_d < 50) %>% 
  mutate(PER = DOC_umolC_g_hr/NPP_umolC_g_hr *100) %>% 
  filter(NPP_umolC_g_hr > 0) 

lm_PER_PAR <- lm(PER ~ PAR, data = PER_PAR)

summary(lm_PER_PAR)



PER_CN10 <- 10*-0.0468 + 3.4496
PER_CN40 <- 40*-0.0468 + 3.4496

```

# Stats: Mature DOC release, & Summary Stats for Mature PER



```{r}

# Filter DOC Release Mature 

MP_DOC_NPP_phys_env_sum_mature <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season %in% c("Spring", "Summer")) %>% 
  filter(age_d <= 50, flag!= 2)


# Mature DOC Release: Spring 

MP_DOC_mature_Spring <- MP_DOC_NPP_phys_env_sum_mature %>% 
  filter(season == "Spring")

# Mature DOC Release: Summer

MP_DOC_mature_Summer <- MP_DOC_NPP_phys_env_sum_mature %>% 
  filter(season == "Summer") 



mature_DOC_NPP_model2 <- lmodel2(DOC_umolC_g_hr~NPP_umolC_g_hr, data = MP_DOC_NPP_phys_env_sum_mature)

mature_DOC_NPP_model2


Spring_mature_DOC_NPP_model2 <- lmodel2(DOC_umolC_g_hr~NPP_umolC_g_hr, data = MP_DOC_mature_Spring)

Summer_mature_DOC_NPP_model2 <- lmodel2(DOC_umolC_g_hr~NPP_umolC_g_hr, data = MP_DOC_mature_Summer)

Spring_mature_DOC_NPP_model2

Summer_mature_DOC_NPP_model2

# Summary Stats: Avg PER when PAR > 0

MP_DOC_NPP_phys_env_sum_mature %>% 
  filter(season != "NA") %>% 
  filter(lightlevel_timepoint != 1) %>% 
  group_by(season) %>% 
  summarize(mean_PER = mean(DOC_umolC_g_hr/NPP_umolC_g_hr),
            sd_PER = sd(DOC_umolC_g_hr/NPP_umolC_g_hr),
            max_per = max(DOC_umolC_g_hr/NPP_umolC_g_hr),
            min_per = min(DOC_umolC_g_hr/NPP_umolC_g_hr)) %>% 
  ungroup()


MP_PER_phys_state <- MP_DOC_NPP_phys_env_sum %>% 
  mutate(phys_state = case_when(
    age_d > 50 ~ "Senescent",
    age_d < 50 ~ "Mature"
  )) %>% 
  filter(season != "NA") %>% 
  filter(lightlevel_timepoint != 1) %>% 
  filter(NPP_umolC_g_hr > 10) %>% 
  group_by(phys_state) %>% 
  summarize(mean_PER = mean(DOC_umolC_g_hr/NPP_umolC_g_hr)*100,
            sd_PER = sd(DOC_umolC_g_hr/NPP_umolC_g_hr)*100,
            max_per = max(DOC_umolC_g_hr/NPP_umolC_g_hr)*100,
            min_per = min(DOC_umolC_g_hr/NPP_umolC_g_hr)*100) %>% 
  ungroup()


MP_DOC_NPP_phys_env_sum_mature %>%
  filter(lightlevel_timepoint != 1) %>% 
  summarize(mean_PER = mean(DOC_umolC_g_hr/NPP_umolC_g_hr),
            sd_PER = sd(DOC_umolC_g_hr/NPP_umolC_g_hr),
            max_per = max(DOC_umolC_g_hr/NPP_umolC_g_hr),
            min_per = min(DOC_umolC_g_hr/NPP_umolC_g_hr)) %>% 
  ungroup()

# test for significant differences in DOC release in MAture kelp between Summer and Spring 


MP_DOC_NPP_PER <- MP_DOC_NPP_phys_env_sum_mature %>% 
  filter(lightlevel_timepoint != 1) %>% 
  filter(season %in% c("Summer", "Spring")) %>% 
  mutate(PER = (DOC_umolC_g_hr / NPP_umolC_g_hr)*100) %>% 
  select(season, PER)


# Boxplot PER Summer Spring Mature 

ggboxplot(MP_DOC_NPP_PER, x = "season", y = "PER", 
          color = "season", palette = c("#00AFBB", "#E7B800"),
          ylab = "PER", xlab = "season")

ggplot(MP_DOC_NPP_PER, aes(x = PER, fill = season, colour = season)) + 
  geom_histogram(alpha = 0.5, position = "identity")

## Test for Normality


with(MP_DOC_NPP_PER, shapiro.test(PER[season == "Summer"])) # p = 0.00014 = not normal

with(MP_DOC_NPP_PER, shapiro.test(PER[season == "Spring"])) # p = 0.01039 = not normal

# Not normal using Wilcox 


season_PER.wilcox <- wilcox.test(PER ~ season, 
                                 data = MP_DOC_NPP_PER)


season_PER.wilcox # p = 0.039 but difference in means only 0.4% 

# Fold change between DOC release in the dark and at saturating irradiances 


avg_DOC_rates_phys_state_light_Dark <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2 & season %in% c("Spring", "Summer")) %>% 
  mutate(phys_state = case_when(
    age_d >= 50 ~ "Senescenct",
    age_d < 50 ~ "Mature")) %>% 
  filter(lightlevel_timepoint %in% c(1,5,6)) %>% 
  mutate(light_level = case_when(lightlevel_timepoint == 1 ~ "Dark",
                   lightlevel_timepoint %in% c(5,6) ~ "Saturating")) %>% 
  group_by(light_level) %>% 
  summarize(mean_DOC = round(mean(DOC_umolC_g_hr),1),
            sd_DOC = round(sd(DOC_umolC_g_hr),1),
            min_PAR = min(PAR),
            max_PAR = max(PAR))

write_excel_csv(avg_DOC_rates_phys_state_light_Dark, "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_DOC/Data/avg_DOC_rates_phys_state_light_Dark.csv")


```

# Mature DOC Release vs PAR 

```{r}


# DOC Production vs. Light 

mature_DOC_PAR_model1 <- lm(DOC_umolC_g_hr~PAR, data = MP_DOC_NPP_phys_env_sum_mature)

summary(mature_DOC_PAR_model1)

doc_PAR_mature <- MP_DOC_NPP_phys_env_sum_mature %>% 
  filter(season %in% c("Summer", "Spring")) %>% 
  ggplot(aes(x = PAR, y = DOC_umolC_g_hr)) +
  geom_point(size = 7) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  labs(color = "PAR") + 
  scale_y_continuous(labels = label_number(accuracy = 1), limits = c(0,10), breaks = c(0,5,10,15,20)) +
  geom_abline(slope = 0.0028, intercept = 2.28, size = 2) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))







```

# Senescent DOC Release vs PAR

```{r}

MP_DOC_NPP_phys_env_sum_senescent <- MP_DOC_NPP_phys_env_sum_flags %>%
  filter(age_d > 50)
  
MP_DOC_NPP_phys_env_sum_senescent %>% 
  ggplot(aes(x = PAR, y = DOC_umolC_g_hr)) +
  geom_point()

lm_DOC_senescence <- lm(DOC_umolC_g_hr ~ PAR, data = MP_DOC_NPP_phys_env_sum_senescent)

summary(lm_DOC_senescence)

MP_DOC_NPP_phys_env_sum_senescent %>% 
  group_by(PAR) %>% 
  summarize(mean_DOC = mean(DOC_umolC_g_hr),
            sd_DOC = sd(DOC_umolC_g_hr))

MP_DOC_NPP_phys_env_sum_senescent <- MP_DOC_NPP_phys_env_sum_flags %>%
  filter(age_d > 50)
  
MP_DOC_NPP_phys_env_sum_senescent %>% 
  ggplot(aes(x = PAR, y = DOC_umolC_g_hr)) +
  geom_point()

lm_DOC_senescence <- lm(DOC_umolC_g_hr ~ PAR, data = MP_DOC_NPP_phys_env_sum_senescent)

summary(lm_DOC_senescence)

MP_DOC_NPP_phys_env_sum_senescent %>% 
  group_by(PAR) %>% 
  summarize(mean_DOC = mean(DOC_umolC_g_hr),
            sd_DOC = sd(DOC_umolC_g_hr))

```



# Physiological Changes


```{r}

Chl_C_age_summer <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season == "Summer") %>% 
  group_by(age_d) %>% 
  filter(Chl_C != "NA", 
         lightlevel_timepoint %in% c(1,2)) %>% 
  summarise(Chl_C_avg = mean(Chl_C), 
            sd_Chl_C = sd(Chl_C),
            season = season) %>% 
  ungroup() 




Chl_C_age_spring <- MP_DOC_NPP_phys_env_sum %>% 
  filter(season == "Spring") %>% 
  group_by(age_d) %>% 
  filter(Chl_C != "NA", 
         lightlevel_timepoint %in% c(1,2)) %>% 
  summarise(Chl_C_avg = mean(Chl_C), 
            sd_Chl_C = sd(Chl_C),
            season = season) %>% 
  ungroup() 



 Chl_C_age <- MP_DOC_NPP_phys_env_sum %>% 
  group_by(age_d) %>% 
  filter(Chl_C != "NA", 
         lightlevel_timepoint %in% c(1,2)) %>% 
  summarise(Chl_C_avg = mean(Chl_C), 
            sd_Chl_C = sd(Chl_C)) %>% 
  ungroup() %>% 
  ggplot(aes(x = age_d, 
             y = Chl_C_avg*1000)) +
  geom_point(size = 6) +
  geom_sigmoid(aes(y = Chl_C_avg[age_d==min(age_d)]*1000, 
                   yend = Chl_C_avg[age_d==max(age_d)]*1000, 
                   x = min(age_d), 
                   xend =max(age_d)), size = 1) +
  ylab(expression(Chl:C~(mg^{"1"} ~ g^{"-1"}))) +
  labs(x = "Age (d)") + 
  scale_y_continuous(labels = label_number(accuracy = 1), 
                     breaks = c(0,2,4,6,8,10), limits = c(0,10)) +
  geom_errorbar(aes(x = age_d, 
                    y = Chl_C_avg*1000, 
                    ymin =  Chl_C_avg*1000 - sd_Chl_C*1000, 
                    ymax = Chl_C_avg*1000 + sd_Chl_C*1000), 
                width = 0) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

 
 
  
Chl_C_age_cohorts <- ggplot() +
  geom_point(aes(x = Chl_C_age_summer$age_d, 
             y = Chl_C_age_summer$Chl_C_avg*1000), 
             shape = 17, 
             size = 5) +
  geom_point(aes(x = Chl_C_age_spring$age_d, 
             y = Chl_C_age_spring$Chl_C_avg*1000),
             shape = 19, 
             size = 5) +
  geom_errorbar(aes(x = Chl_C_age_spring$age_d, 
                    y = Chl_C_age_spring$Chl_C_avg*1000, 
                    ymin =  Chl_C_age_spring$Chl_C_avg*1000 - Chl_C_age_spring$sd_Chl_C*1000, 
                    ymax = Chl_C_age_spring$Chl_C_avg*1000 + Chl_C_age_spring$sd_Chl_C*1000), 
                width = 0) +
  geom_sigmoid(aes(y = Chl_C_age_summer$Chl_C_avg[Chl_C_age_summer$age_d==min(Chl_C_age_summer$age_d)]*1000, 
                   yend = Chl_C_age_summer$Chl_C_avg[Chl_C_age_summer$age_d==max(Chl_C_age_summer$age_d)]*1000, 
                   x = min(Chl_C_age_summer$age_d), 
                   xend = 100), 
               size = 1) +
  ylab(expression(Chl:C~(mg^{"1"} ~ g^{"-1"}))) +
  labs(x = "Age (d)") + 
  geom_errorbar(aes(x = Chl_C_age_summer$age_d, 
                    y = Chl_C_age_summer$Chl_C_avg*1000, 
                    ymin =  Chl_C_age_summer$Chl_C_avg*1000 - Chl_C_age_summer$sd_Chl_C*1000, 
                    ymax = Chl_C_age_summer$Chl_C_avg*1000 + Chl_C_age_summer$sd_Chl_C*1000), 
                width = 0) +
  geom_sigmoid(aes(y = Chl_C_age_spring$Chl_C_avg[Chl_C_age_spring$age_d==min(Chl_C_age_spring$age_d)]*1000, 
                   yend = Chl_C_age_summer$Chl_C_avg[Chl_C_age_summer$age_d==max(Chl_C_age_summer$age_d)]*1000, 
                   x = min(Chl_C_age_spring$age_d), 
                   xend = 100), size = 1, shape = 16) +
  geom_vline(xintercept = 50, linetype = "dashed", size = 0.5) +
  ylab(expression(Chl:C~(mg^{"1"} ~ g^{"-1"}))) +
  labs(x = "Age (d)") + 
  scale_y_continuous(labels = label_number(accuracy = 1), 
                     breaks = c(0,5,10,15,20), 
                     limits = c(0,20)) +
  scale_shape_manual(labels = c("Spring", "Summer"), values = c(19,17), guide = ) +
  labs(color = "",
       shape = "Season") +
  coord_cartesian(expand = c(0,0)) +
  scale_x_continuous(breaks = c(25,50,75,100), limits = c(10,100)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = c(0.8,0.8),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times")) +
   annotate("rect", xmin = 10, xmax = 50, ymin = 0, ymax = 20,
           alpha = .25,fill = "#266A25")  +
  annotate("rect", xmin = 50.1, xmax = 100, ymin = 0, ymax = 20,
           alpha = .25,fill = "#BD9C3D") 

  




```


# Tissue Nitrogen Both Cohorts


```{r}

# Tissue Nitrogen

tissue_N_age_both_seasons <- MP_DOC_NPP_phys_env_sum %>% 
  group_by(age_d, season) %>% 
  filter(chlA_g_g != "NA", 
         lightlevel_timepoint %in% c(1,2),
         reef == "MK",
         season %in% c("Summer", "Spring")) %>% 
 summarise(N_avg = mean(tissue_N_percent), 
            sd_N = sd(tissue_N_percent)) %>%  
  ungroup() %>% 
  ggplot(aes(x = age_d, 
             y = N_avg,
             color = season)) +
  geom_point(size = 6) +
  geom_line(aes(color = season), size = 1) +
  labs(x = "Age (d)",
       y = "Tissue Nitrogen (% Dry Weight)") + 
  scale_color_manual(values = c("red", "black")) +
  scale_y_log10() +
  geom_errorbar(aes(x = age_d, 
                    y = N_avg, 
                    ymin =  N_avg - sd_N, 
                    ymax = N_avg + sd_N),
                width = 0) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

  




```



# Stats: Differnces in Chl:C and Tissue C:N between cohorts and physiological status (mature vs. senescent)

```{r}

MP_chlC_sum <- MP_DOC_NPP_phys_env_sum %>% 
  group_by(age_d, season) %>% 
  filter(chlA_g_g != "NA", 
         lightlevel_timepoint %in% c(1,2),
         reef == "MK",
         season %in% c("Summer", "Spring")) %>% 
  select(blade, age_d, season, Chl_C, C_N, tissue_N_percent, chlA_nmol_g) %>% 
  mutate(physio_state = case_when(
    age_d <= 50 ~ "Mature",
    age_d >50 ~ "Senescent"
  ))

# Summarize ChlC and CN of mature and senescent blades from Spring and Summer cohorts

MP_ChlC_CN_summary_statistics <- MP_chlC_sum %>% 
  group_by(season, physio_state) %>% 
  summarize(mean_Chl_C = mean(Chl_C)*1000,
            sd_Chl_C = sd(Chl_C)*1000,
            mean_CN = mean(C_N),
            sd_CN = sd(C_N),
            mean_N = mean(tissue_N_percent),
            sd_N = sd(tissue_N_percent),
            mean_Chl = mean(chlA_nmol_g),
            sd_Chl = sd(chlA_nmol_g))

# Two Way ANOVA: Chl:C vs Season * Physiological State

ggboxplot(MP_chlC_sum, x = "season", y = "Chl_C", color = "physio_state",
          palette = c("#00AFBB", "#E7B800"))


ChlC_physiostate_season.aov2 <- aov(Chl_C ~ season + physio_state, data = MP_chlC_sum)



summary(ChlC_physiostate_season.aov2)


# Two Way Interaction between Season and Physiostate

ChlC_physiostate_season.aov3 <- aov(Chl_C ~  season + physio_state, data = MP_chlC_sum)


summary(ChlC_physiostate_season.aov3) # No significant interaction 

# Pairwise TUKEY HSD Test

TukeyHSD(ChlC_physiostate_season.aov3)

pairWise_comparison_ChlC_season_physio_state # Significant difference in Chl:C between mature kelp between seasons and between mature and senescent kelp within seasons.



CN_physiostate_season.aov3 <- aov(C_N ~  season + physio_state, data = MP_chlC_sum)

CN_physiostate_season.aov3

pairwise_comparison_CN_season_physio_state <- TukeyHSD(CN_physiostate_season.aov3)

pairwise_comparison_CN_season_physio_state

# Measure percent N between phys state and season 

tissueN_physiostate_season.aov3 <- aov(tissue_N_percent ~  season * physio_state, data = MP_chlC_sum)

summary(tissueN_physiostate_season.aov3)

pairwise_comparison_tissueNpercent_season_physio_state <- TukeyHSD(tissueN_physiostate_season.aov3)

pairwise_comparison_tissueNpercent_season_physio_state


# Two Sample t-test: C:N content of mature kelp in Spring vs. Summer


## Test for Normality

mature_CN <- MP_chlC_sum %>% 
  filter(physio_state == "Mature") 

with(mature_CN, shapiro.test(C_N[season == "Summer"])) # p = 0.2471 = normal

with(mature_CN, shapiro.test(C_N[season == "Spring"])) # p = 0.7502 = normal

## Equal Variances 

matureCN.ftest <- var.test(C_N ~ season, data = mature_CN)

matureCN.ftest # Variances are not equal. Use Welch's t-test 


summer_mature_CN <- mature_CN %>% 
  filter(season == "Summer") 

summer_mature_CN_vector <- summer_mature_CN$C_N

spring_mature_CN <- mature_CN%>% 
  filter(season == "Spring") 

spring_mature_CN_vector <- spring_mature_CN$C_N


mature_seasonal_CN_t.test <- t.test(C_N ~ season, data = mature_CN, var.equal = FALSE, alternative = "less")

mature_seasonal_CN_t.test # Spring time CN is significantly lower than Summer time CN for mature kelp


.chl_tissueN_lm <- lm(data = MP_chlC_sum, chlA_nmol_g ~ tissue_N_percent)

summary(chl_tissueN_lm)

MP_chlC_sum %>% 
  ggplot(aes(x = tissue_N_percent, y = chlA_nmol_g)) +
  geom_point()



# Two Sample t-test: C:N content of mature kelp in Spring vs. Summer


## Test for Normality

mature_ChlC <- MP_chlC_sum %>% 
  filter(physio_state == "Mature") 

with(mature_ChlC, shapiro.test(Chl_C[season == "Summer"])) # p = 0.0.003 = not normal

with(mature_ChlC, shapiro.test(Chl_C[season == "Spring"])) # p = 0.266 = normal

## Equal Variances 

matureChlC.ftest <- var.test(Chl_C ~ season, data = mature_ChlC)

matureChlC.ftest # Variances are not equal. Use Welch's t-test 




mature_seasonal_ChlC_wilcox_test <- wilcox.test(Chl_C ~ season, data = mature_ChlC, exact = FALSE)

mature_seasonal_ChlC_wilcox_test # Spring time CN is significantly lower than Summer time CN for mature kelp




```

# Boxplots: Tissue Nitrogen between season and physiological state 

```{r}

season_phys_state_boxplot <- MP_chlC_sum %>% 
  ggplot() +
  geom_boxplot(aes(x = season, y = tissue_N_percent, fill = physio_state)) +
  geom_point(aes(x = season, y = tissue_N_percent, color = physio_state),
             position = position_jitterdodge()) +
  labs(y = "Tissue Nitrogen (% DW)",
       x = "") +
  scale_fill_manual(values = c("#266A25","#BD9C3D")) +
  scale_color_manual(values = c("darkgreen","darkgoldenrod")) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = c(0.72,0.75),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

  




```



# Stats: Linear Regression between Chl:C and Tissue CN


```{r}


MP_chlC_sum %>% 
  ggplot(aes(x = C_N, y = Chl_C)) +
  geom_point(size = 5,aes(color = season)) +
  geom_smooth(method = "lm", aes(color = season)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

# Multiple Linear Effects Model: Chl ~ CN + Season

mle_lm_ChlC_CN_season <- lm(Chl_C ~ C_N + season, data = MP_chlC_sum)

summary(mle_lm_ChlC_CN_season)

# Linear Model 
lm_ChlC_CN <-  lm(Chl_C ~ C_N, data = MP_chlC_sum)

summary(lm_ChlC_CN)

AIC(mle_lm_ChlC_CN_season, lm_ChlC_CN)


```


# NPP 


```{r}


NPP_data <- read_xlsx("MP_NPP_summary.xlsx", sheet = 2)

NPP_PAR <- NPP_data %>% 
  group_by(age_d) %>% 
  ggplot(aes(x = PAR, y = modeled_NPP, color = age_d, group = age_d)) +
  geom_path(size = 4, orientation = "y") +
  xlab(expression(PAR~(µmol ~ m^{"-2"} ~ s^{"-1"}))) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "Age (d)") + 
  scale_y_continuous(labels = label_number(accuracy = 1)) +
  scale_color_viridis_b(breaks = c(0,15,30,45,60,75,90), limits = c(0,90)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))






```


# NPP MLR w/ Light and Chl:C


```{r}


MP_DOC_NPP_phys_env_sum_noflags  <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2)


MP_DOC_NPP_phys_env_sum_noflags %>% 
  ggplot(aes(x = Chl_C, y = NPP_umolC_g_hr)) +
  geom_point(aes(color = PAR), size = 7) +
  ylab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(Chl:C~(g^{"1"} ~ g^{"-1"}))) +
  labs(color = "PAR") + 
  scale_y_continuous() +
  scale_x_continuous(labels = label_number(accuracy = 0.001), breaks = c(0, 0.004, 0.008, 00.012)) +
  scale_color_viridis_b(breaks = c(seq(0,1500,250)), limits = c(0,1500)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))





mlr_NPP_PAR_ChlC <- lm(mean_NPP ~ mean_PAR + mean_Chl_C, data = MP_DOC_NPP_phys_env_sum_noflags)



summary(mlr_NPP_PAR_ChlC)

plot(mlr_NPP_PAR_ChlC)

mlr_NPP_PAR_ChlC


```



# Generalized Additive Model of NPP ~ Chl:C and PAR



```{r}


gam_NPP_PAR_age_d <- gam(NPP_umolC_g_hr ~ s(age_d) + s(PAR, k = 1),data = MP_DOC_NPP_phys_env_sum_noflags)



summary(gam_NPP_PAR_age_d)


plot(gam_NPP_PAR_age_d, pch = 1, cex = 1, residuals = T, pages = 1, ylab = "Additive Effect on NPP")

gam.check(gam_NPP_PAR_ChlC)


# Extract fitted values and measured to visualize model fit 

gam_x <- gam_NPP_PAR_ChlC$fitted.values
gam_y <- gam_NPP_PAR_ChlC$y

gam_model_fit_df <- data_frame(gam_x,gam_y)

gam_fit_model <- gam_model_fit_df %>% 
  ggplot(aes(gam_NPP_PAR_ChlC$fitted.values, gam_NPP_PAR_ChlC$y)) + 
  geom_point() +
  geom_abline(slope = 1.02, intercept = -1.5) +
  geom_abline(color = "red") +
  xlab(expression(Fitted~NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  ylab(expression(Measured~NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  scale_y_continuous(limits = c(-50,250), breaks = c(-50,0,50,100,150,200,250)) +
  scale_x_continuous(limits = c(-50,250), breaks = c(-50,0,50,100,150,200,250)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

gam_NPP_fit_regression <- lm(gam_y~gam_x, data = gam_model_fit_df)

summary(gam_NPP_fit_regression)


AIC(gam_NPP_PAR_ChlC_1, gam_NPP_PAR_ChlC_2)
```

# GAM FOR DOC PRODUCTION

# Generalized Additive Model of DOC ~ Chl:C and PAR

```{r}

gam_DOC_PAR_ChlC <- gam(DOC_umolC_g_hr ~ s(Chl_C, k = 2) + s(PAR) ,data = MP_DOC_NPP_phys_env_sum_noflags)

summary(gam_DOC_PAR_ChlC)


plot(gam_DOC_PAR_ChlC, pch = 1, cex = 1, residuals = T, pages = 1, ylab = "Additive Effect on DOC")

gam.check(gam_DOC_PAR_ChlC_1)

# Extract fitted values and measured to visualize model fit 

gam_DOC_x <- gam_DOC_PAR_ChlC$fitted.values
gam_DOC_y <- gam_DOC_PAR_ChlC$y

gam_DOC_model_fit_df <- data_frame(gam_DOC_x,gam_DOC_y)

gam_DOC_fit_model <- gam_DOC_model_fit_df %>% 
  ggplot(aes(gam_DOC_x, gam_DOC_y)) + 
  geom_point() +
  geom_abline(slope = 1.02, intercept = -1.5 ,color = "red") +
  xlab(expression(Fitted~DOC~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  ylab(expression(Measured~DOC~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  scale_y_continuous(limits = c(0,70), breaks = c(0,10,20,30,40,50,60,70)) +
  scale_x_continuous(limits = c(0,70), breaks = c(0,10,20,30,40,50,60,70)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

gam_DOC_fit_regression <- lm(gam_DOC_y~gam_DOC_x, data = gam_DOC_model_fit_df)

summary(gam_NPP_fit_regression)


```



# DOM Composition 



```{r}

MP_comp_sum_TCHO <- MP_DOM_Sugar %>% 
  group_by(age_d) %>% 
  summarize(HPLC_TCHO = mean(HPLC_sugar_per_doc), sd_HPLC_TCHO = sd(HPLC_sugar_per_doc),
            TPC = mean(a270_DOC), sd_TPC = sd(a270_DOC),
            mean_fucose = mean(Fucose),
            sd_fucose = sd(Fucose),
            mean_mannuronic = mean(`Man-URA`),
            sd_mannuronic = sd(`Man-URA`))
  

TCHO_DOC <- MP_comp_sum_TCHO %>% 
   ggplot(aes(x = age_d, y = HPLC_TCHO)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = HPLC_TCHO - sd_HPLC_TCHO, ymax = HPLC_TCHO + sd_HPLC_TCHO)) %>% 
  labs(x = "Age (d)",
       y = "HPLC TCHO (%DOC)") + 
  scale_y_continuous(labels = label_number(accuracy = 1), limits = c(0,50), breaks = c(0,10,20,30,40,50)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

  

TPC_DOC <- MP_comp_sum_TCHO %>% 
  ggplot(aes(x = age_d, y = TPC*100)) +
  geom_point(size = 5) +
  labs(x = "Age (d)",
       y = "TPC (%DOC)") + 
  scale_y_continuous(labels = label_number(accuracy = 1), limits = c(0,50), breaks = c(0,10,20,30,40,50)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))







```



```{r}


percent_CHO <- MP_DOM_character %>% 
  mutate(age_d = as.character(age_d)) %>% 
  ggplot(aes(x = age_d, y = HPLC_sugar_per_doc)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "Percent CHO (%)",
       fill = "") +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))

mean_CHO <- MP_DOM_character %>% 
  summarize(mean(HPLC_sugar_per_doc))

# ANOVA 



model  <- lm( HPLC_sugar_per_doc ~ age_d, data = MP_DOM_character)
ggqqplot(residuals(model))
shapiro_test(residuals(model))

MP_DOM_character %>% 
  group_by(age_d) %>%
  identify_outliers(HPLC_sugar_per_doc)


ggqqplot(MP_DOM_character, "HPLC_sugar_per_doc", facet.by = "age_d")

MP_DOM_character %>% 
  group_by(age_d) %>% 
  shapiro_test(HPLC_sugar_per_doc)


# Filter out extreme outlier to compare ANOVA test with and without it 

MP_DOM_character_no_outlier <- MP_DOM_character %>% 
  filter(blade != 2 | age_d != 16)

# Anova 

res.aov <- MP_DOM_character %>% anova_test(HPLC_sugar_per_doc ~ age_d)
res.aov

res.aov_no_extreme_outlier <- MP_DOM_character_no_outlier %>% anova_test(HPLC_sugar_per_doc ~ age_d)
res.aov_no_extreme_outlier

pwc <- MP_DOM_character %>% tukey_hsd(HPLC_sugar_per_doc ~ as.factor(age_d))
pwc

pwc_no_outlier <- MP_DOM_character_no_outlier %>% tukey_hsd(HPLC_sugar_per_doc ~ as.factor(age_d))
pwc_no_outlier
```


# Sugar Monomers Composition & PCA


```{r}

suga_numeric <- sugars[1:42,7:15]



```


```{r}


sug.pca1 <- prcomp(suga_numeric, scale = TRUE)

fviz_eig(sug.pca1)

sug.pca1

summary(sug.pca1)

get_eig(sug.pca1)

eigenvals(sug.pca1)

print(sug.pca1)



```


 # Sugars Biplot for Mature vs. Senescent Kelp Exudates


```{r}

library(scico)

fviz_pca_biplot(sug.pca1,
             col.var = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             col.ind = "black",
             repel = TRUE     # Avoid text overlapping
             )


pca_plot <- fviz_pca_biplot(sug.pca1, 
                level = "var",
                habillage = sugars[1:42,]$phase, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("#266A25","#BD9C3D")) +
  labs(title = "", 
       x = "PC1 (24%)",
       y = "PC2 (24%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = "none",
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

# PCA Plot Season 

pca_plot_season <- fviz_pca_biplot(sug.pca1, 
                level = "var",
                habillage = MP_DOM_Sugar[1:42,]$season, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("#022E73", "#9D3709")) +
  labs(title = "", 
       x = "PC1 (26%)",
       y = "PC2 (23%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = c(0.2,0.1),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
  
# Compare Mature Kelp for Summer and Spring

fviz_pca_biplot(sug.pca_mature,
             col.var = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             col.ind = "black",
             repel = TRUE     # Avoid text overlapping
             )


pca_plot_mature <- fviz_pca_biplot(sug.pca_mature, 
                level = "var",
                habillage = MP_DOM_Sugar_mature[1:12,]$season, 
                addEllipses = TRUE, 
                ellipse.level = 0.95,
                label = "var", labelsize = 5,
                col.var = "black",
                repel = T,
                geom="point", pointsize = 4,
                palette = c("steelblue1", "firebrick1")) +
  labs(title = "", 
       x = "PC1 (26%)",
       y = "PC2 (23%)") +
  scale_y_continuous(labels = label_number(accuracy = 1)) + 
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.position = c(0.15,0.10),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
  





```


# Permanova for Sugar Mol% between Mature and Senescent phase Kelp


```{r}

MP_DOM_sugar_c <- scale(MP_DOM_Sugar[1:42,6:14])

# PERMANOVA of differences in carbohydrate composition by physiological state

MP_DOM_sugar_physstate_adonis <- adonis(MP_DOM_sugar_c~phase, data = MP_DOM_Sugar[1:42,], method = "eu")

MP_DOM_sugar_physstate_adonis # R2 = 0.14 p = 0.001


# PERMANOVA of differences in carbohydrate composition by season 

MP_DOM_sugar_season_adonis <- adonis(MP_DOM_sugar_c~season, data = MP_DOM_Sugar[1:42,], method = "eu")

MP_DOM_sugar_season_adonis # R2 = 0.07 p = 0.007 **

MP_DOM_sugar_phys_state_season_adonis <- adonis(MP_DOM_sugar_c~ phase + season, data = MP_DOM_Sugar[1:42,], method = "eu")

MP_DOM_sugar_phys_state_season_adonis

pairwiseAdonis::pairwise.adonis2(MP_DOM_sugar_phys_state_season_adonis)

```


# Fucose & Mannuronic Acid Boxplot over Time: SUMMER


```{r}

fucose_manURA_age <- MP_DOM_Sugar %>% 
  filter(season == "Summer") %>% 
  select(age_d, Fucose, `Man-URA`, blade) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent") %>% 
  mutate(age_d = as.factor(age_d))
  
boxplot_fucose_manURA_summer <- fucose_manURA_age %>% 
  ggplot(aes(x = age_d, y = mol_percent*100, fill = name)) +
  geom_boxplot() +
   ggtitle(label = "Summer") + 
  labs(x = "Age (d)",
       y = "Mol%",
       fill = "") +
  scale_fill_manual(values = c("#022E73", "#9D3709")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        plot.title = element_text(size = 24),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.position = "none",
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))




```


# Fucose & Mannuronic Acid Boxplot over Time: SPRING


```{r}

fucose_manURA_age_spring <- MP_DOM_Sugar %>% 
  filter(season == "Spring") %>% 
  select(age_d, Fucose, `Man-URA`, blade) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent") %>% 
  mutate(age_d = as.factor(age_d))
  
boxplot_fucose_manURA_spring <- fucose_manURA_age_spring %>% 
  ggplot(aes(x = age_d, y = mol_percent*100, fill = name)) +
  geom_boxplot() +
  labs(x = "Age (d)",
       y = "",
       fill = "") +
  ggtitle(label = "Spring") + 
  scale_fill_manual(values = c("#022E73", "#9D3709")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        plot.title = element_text(size = 24), 
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))




```

# Plot Avg DOC Release Rate by Blade against mol% of Mannuronic Acid

```{r}
DOC_mean_by_age_season_blade <- MP_DOC_NPP_phys_env_sum %>% 
  filter(flag != 2 & season %in% c("Spring", "Summer")) %>% 
  group_by(age_d, blade, season) %>% 
  summarize(mean_DOC = mean(DOC_umolC_g_hr),
            sd_DOC = sd(DOC_umolC_g_hr))

DOC_Sugar <- merge(DOC_mean_by_age_season_blade, MP_DOM_Sugar, by = c("age_d", "blade", "season"))


DOC_ManURA_p <- DOC_Sugar %>%
  ggplot(aes(x = `Man-URA`*100, y = mean_DOC)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_DOC - sd_DOC, ymax = mean_DOC + sd_DOC), width = 0) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(x = "Man-URA (mol%)") +
  geom_abline(intercept = -6.22, slope = 89.67/100, linetype = "dashed", color = "black") +
 theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

# DOC Release Rate ManURA LM


DOC_ManURA_lm = lm(data = DOC_Sugar, mean_DOC ~ `Man-URA`)

summary(DOC_ManURA_lm)# R2 = 0.48, p = 1.947e-07

```






# Summary: Avg Sugar mol% in Mature vs. Senescent phase

```{r}

sugar_summary <- MP_DOM_Sugar %>% 
  select(phase, season,Fucose:`Man-URA`) %>% 
  pivot_longer(cols = Fucose: `Man-URA`,
               names_prefix = "sugar",
               values_to = "mol_percent")%>% 
  group_by(phase, name, season) %>% 
  summarize(mean_molp = mean(mol_percent),
            sd_molp = sd(mol_percent))
  
sugars_fold_change_phys_state <- sugar_summary %>% 
  group_by(name) %>% 
  summarise(fold_change = mean_molp[phase == "Senescent"]/mean_molp[phase == "Mature"])

sugars_fold_change_season <- sugar_summary %>% 
  group_by(phase, name) %>% 
  summarise(fold_change = mean_molp[season == "Spring"]/mean_molp[season == "Summer"])

```


# Glucosamine Mature (16 days) Box Plot between Spring and Summer


```{r}

glucosamine_season_mature <- sugars %>% 
  filter(age_d < 50) %>% 
  select(season, Glucoseamine, blade) %>% 
  pivot_longer(cols = Glucoseamine,
               names_prefix = "sugar",
               values_to = "mol_percent") 
  
boxplot_glucosamine_season_mature <- glucosamine_season_mature %>% 
  ggplot(aes(x = season, y = mol_percent*100)) +
  geom_boxplot(fill = "grey") +
  labs(x = "",
       y = "Mol%",
       fill = "") +
  scale_fill_manual(values = c("grey")) +
  geom_point( position=position_dodge(width=0.75)) +
   theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))



```

# Glucosamine age and season 

```{r}

MP_DOM_Sugar_glucosamine <- sugars %>% 
  filter(age_d %in% c(16, 77, 78)) %>% 
  select(age_d, season, blade, Glucoseamine) %>% 
  mutate(phase = case_when(
    age_d >= 50 ~ "Senescent",
    age_d < 50 ~ "Mature"
  ))


glucosamine_season <- MP_DOM_Sugar_glucosamine %>% 
  select(age_d, season, Glucoseamine, blade, phase) %>% 
  pivot_longer(cols = Glucoseamine,
               names_prefix = "sugar",
               values_to = "mol_percent") 


glucosamine_molp_boxplot <- glucosamine_season %>% 
  ggplot() +
  geom_boxplot(aes(x = season, y = mol_percent*100, fill = phase)) +
  geom_point(aes(x = season, y = mol_percent*100, color = phase),
             position = position_jitterdodge(0.1)) +
  labs(y = "Glucosamine (mole%)",
       x = "",
       color = "",
       fill = "") +
  scale_fill_manual(values = c("#266A25","#BD9C3D")) +
  scale_color_manual(values = c("darkgreen","darkgoldenrod")) +
  scale_y_continuous(limits = c(0,20)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

# Two-way ANOVA Glucosamine vs. Season and Physiology


aov.glucosamine.season.phase <- aov(data = MP_DOM_Sugar_glucosamine, Glucoseamine ~ season + phase)

summary(aov.glucosamine.season.phase)

TukeyHSD(aov.glucosamine.season.phase)


```


# Gluconic acid age and season 


```{r}

MP_DOM_Sugar_glcURA <- MP_DOM_Sugar %>% 
  filter(age_d %in% c(16, 77, 78)) %>% 
  select(age_d, season, blade, `Glc-URA`) 


glcURA_season <- MP_DOM_Sugar_glcURA %>% 
  select(age_d, season, `Glc-URA`, blade) %>% 
  pivot_longer(cols = 'Glc-URA' ,
               names_prefix = "sugar",
               values_to = "mol_percent") 


glcURA_molp_boxplot <- glcURA_season %>% 
  mutate(phase = case_when(age_d < 50 ~ "Mature",
                           age_d > 50 ~ "Senescent")) %>% 
   ggplot(aes(phase, mol_percent)) +
  geom_boxplot(position = "dodge") +
  geom_point(size = 3) +
  facet_wrap(~ season, strip.position = "bottom") +
  theme_minimal() +
   theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.placement = "outside",
        strip.background = element_rect(fill = "white", color = "white"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"),
        strip.text = element_text(size = 20, face = "bold")) +
  labs(x = "",
       y = "Glc-URA (Mol%)")



```


# Repeated Measures ANOVA Fucose/Man-URA over Age: Summer


```{r}

# Outliers


fucose_manURA_age %>% 
  group_by(name, age_d) %>% 
  identify_outliers(mol_percent)

# Normality

fucose_manURA_age %>% 
  group_by(name, age_d) %>% 
  shapiro_test(mol_percent)

# QQ Plot

ggqqplot(fucose_manURA_age, "mol_percent", ggtheme = theme_bw()) +
  facet_grid(age_d ~ name, labeller = "label_both")

# Test

sugar_fucose_manURA_age.aov <- anova_test(
  data = fucose_manURA_age, dv = mol_percent, wid = blade,
  within = c(name, age_d)
  )
get_anova_table(sugar_fucose_manURA_age.aov)


# Effect of treatment at each time point


one.way <- fucose_manURA_age %>%
  group_by(age_d) %>%
  anova_test(dv = mol_percent, wid = blade, within = name) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")
one.way


# Pairwise comparisons between treatment groups


pwc <- fucose_manURA_age %>%
  group_by(age_d) %>%
  pairwise_t_test(
    mol_percent ~ name, paired = TRUE,
    p.adjust.method = "bonferroni"
    )
pwc

write_csv(as.data.frame(pwc), "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_DOC/Data/summer_fucose_manURA_pairwise_t_test.csv")

# Effect of time at each level of treatment

one.way2 <- fucose_manURA_age %>%
  group_by(name) %>%
  anova_test(dv = mol_percent, wid = blade, within = age_d) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

one.way2

# Pairwise comparisons between time points

pwc2 <- fucose_manURA_age %>%
  group_by(name) %>%
  pairwise_t_test(
    mol_percent ~ age_d, paired = TRUE,
    p.adjust.method = "bonferroni"
    )

pwc2

pwc2_csv <- write_csv(as.data.frame(pwc2), "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_DOC/Data/summer_fucose_manURA_pairwise_t_test.csv" )

# Plot 

pwc <- pwc %>% add_xy_position(x = "Age (d)")

bxp <- ggboxplot(
  fucose_manURA_age, x = "age_d", y = "mol_percent", 
  color = "name", palette = c("#00AFBB", "#E7B800")
  )

bxp +
  stat_pvalue_manual(one.way$p.adj, tip.length = 0, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(sugar_fucose_manURA_age.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
  )
  

```

# Repeated Measures ANOVA Fucose/Man-URA over Age: Spring


```{r}

# Outliers


fucose_manURA_age_spring %>% 
  group_by(name, age_d) %>% 
  identify_outliers(mol_percent)

# Normality

fucose_manURA_age_spring %>% 
  group_by(name, age_d) %>% 
  shapiro_test(mol_percent)

# QQ Plot

ggqqplot(fucose_manURA_age_spring, "mol_percent", ggtheme = theme_bw()) +
  facet_grid(age_d ~ name, labeller = "label_both")

# Test

sugar_fucose_manURA_age_spring.aov <- anova_test(
  data = fucose_manURA_age_spring, dv = mol_percent, wid = blade,
  within = c(name, age_d)
  )
get_anova_table(sugar_fucose_manURA_age_spring.aov)


# Effect of treatment at each time point


one.way_spring <- fucose_manURA_age_spring %>%
  group_by(age_d) %>%
  anova_test(dv = mol_percent, wid = blade, within = name) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

one.way_spring


# Pairwise comparisons between treatment groups


pwc_spring <- fucose_manURA_age_spring %>%
  group_by(age_d) %>%
  pairwise_t_test(
    mol_percent ~ name, paired = TRUE,
    p.adjust.method = "bonferroni"
    )

pwc_spring

write_csv(pwc_spring,  "/Users/chanceenglish/Desktop/Lab Shiz/ARPA-e/Giant_Kelp_DOC/Data/spring_fucose_manURA_pairwise_t_test.csv")

# Effect of time at each level of treatment

one.way2_spring <- fucose_manURA_age %>%
  group_by(name) %>%
  anova_test(dv = mol_percent, wid = blade, within = age_d) %>%
  get_anova_table() %>%
  adjust_pvalue(method = "bonferroni")

one.way2_spring

# Pairwise comparisons between time points

pwc2_spring <- fucose_manURA_age_spring %>%
  group_by(name) %>%
  pairwise_t_test(
    mol_percent ~ age_d, paired = TRUE,
    p.adjust.method = "bonferroni"
    )

pwc2_spring

# Plot 

pwc_spring <- pwc_spring %>% add_xy_position(x = "Age (d)")

bxp_spring <- ggboxplot(
  fucose_manURA_age_spring, x = "age_d", y = "mol_percent", 
  color = "name", palette = c("#00AFBB", "#E7B800")
  )

bxp_spring +
  stat_pvalue_manual(one.way_spring$p.adj, tip.length = 0, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(sugar_fucose_manURA_age_spring.aov, detailed = TRUE),
    caption = get_pwc_label(pwc_spring)
  )
  

```

# DOC Flux Senescence Rate and NLS Model DOC vs. ChlA


```{r}

senescent_kelp <- MP_DOC_NPP_phys_env_sum %>% 
  group_by(season) %>%
  filter(flag != 2) %>% 
  summarise(norm_chlA = (chlA_nmol_g/mean(chlA_nmol_g[age_d == min(age_d)]))*100,
            DOC_umolC_g_hr = DOC_umolC_g_hr,
            age_d = age_d,
            PAR = PAR,
            NPP_umolC_g_hr = NPP_umolC_g_hr,
            tissue_C_percent = tissue_C_percent,
            dry_mass_g = dry_mass_g,
            Chl_C = Chl_C,
            blade = blade) %>% 
  filter(age_d > 50)


# NLS Model DOC production rate vs. Normalized Chl concentration 


# Select an approximate $\theta$, since theta must be lower than min(y), and greater than zero
theta.chlA <- min(senescent_kelp$DOC_umolC_g_hr) *0.99 

# Estimate the rest parameters using a linear model
model.chlA <- lm(log(DOC_umolC_g_hr - theta.chlA) ~ norm_chlA, data=senescent_kelp)

alpha.chlA <- exp(coef(model.chlA)[1])
beta.chlA <- coef(model.chlA)[2]

# Starting parameters
start <- list(alpha = alpha.chlA, beta = beta.chlA, theta = theta.chlA)
start

model_nls_senescent_DOC<- nls(DOC_umolC_g_hr ~ alpha * exp(beta * norm_chlA) + theta , data = senescent_kelp, start = start)

model_nls_senescent_DOC

summary(model_nls_senescent_DOC)

R2.nls
# Calculate R2 



# Plot DOC release rate as a function of Chl:C w/ NLS Exponential Model Fit


DOC_normCHLA_p <- senescent_kelp %>% 
  ggplot(aes(x = norm_chlA, y = DOC_umolC_g_hr)) +
   geom_point(size = 5, aes(color = age_d)) +
   geom_smooth(aes(x = senescent_kelp$norm_chlA, y = senescent_kelp$DOC_umolC_g_hr),
    method = "nls", se = FALSE,
    formula = y ~ a * exp(b * x),
    method.args = list(start = c(a = 31.43, b = -0.0331)), color = "black") +
   scico::scale_color_scico(palette = "vik", breaks = c(50,60,70,80), limits = c(50,80)) +
  labs(x = "Normalized Chl (% Initial)",
       color = "Age (d)") +
  scale_x_continuous(breaks = c(0,25,50,75,100), limits = c(0,100)) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))


solubilization_rates <- MP_DOC_NPP_phys_env_sum %>% 
  filter(age_d > 50 & flag != 2) %>% 
  group_by(age_d, blade, season) %>% 
  summarize(DOC = mean(DOC_umolC_g_hr),
            sd_DOC = sd(DOC_umolC_g_hr),
            dry_mass_g = dry_mass_g,
            tissue_C_percent = tissue_C_percent,
            Chl_C = Chl_C*1000,
            daily_solubilization = ((DOC /1000/1000*12)*24)/((tissue_C_percent/100))*100) %>% 
  distinct()
        





solubilization_rate_p <- senescent_kelp %>% 
  mutate(daily_solubilization = ((DOC_umolC_g_hr/1000/1000*12.01)*24)/((tissue_C_percent/100))*100) %>% 
  ggplot(aes(x = Chl_C, y = daily_solubilization/100)) +
  geom_point(size = 5) +
  xlab(expression(Chl:C~(mg^{"1"} ~ g^{"-1"}))) +
  ylab(expression(Solubilization~Rate~(d^{"-1"}))) +
  scale_x_continuous(breaks = c(0,2,4,6,8), limits = c(0,7)) +
  scale_y_continuous(limits = c(0,0.08), breaks = c(0,0.02,0.04,0.06, 0.08)) +
  geom_hline(yintercept = 0.025, linetype = "dashed", size = 1) +
  geom_smooth(method="lm", formula= (y ~ exp(-x)), color=1) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))

solubilization_rates <- senescent_kelp %>% 
  group_by(age_d, blade,season) %>% 
  summarize(DOC_umolC_g_hr = mean(DOC_umolC_g_hr),
            tissue_C_percent = max(tissue_C_percent),
            Chl_C = max(Chl_C),
            dry_mass_g = max(dry_mass_g)) %>% 
  mutate(daily_solubilization = ((DOC_umolC_g_hr/1000/1000*12.01)*24)/((tissue_C_percent/100))*100)


solubilization_plot <- solubilization_rates %>% 
  ungroup() %>% 
  mutate(predicted_solub_rate = predict(solub_chlC_lm, solubilization_rates)) %>% 
  ggplot(aes(x = Chl_C, y = daily_solubilization)) +
  geom_point(size = 5) +
  xlab(expression(Chl:C~(mg^{"1"} ~ g^{"-1"}))) +
  ylab(expression(Solubilization~Rate~("%"~d^{"-1"}))) +
  scale_x_continuous(breaks = c(0,2,4,6,8,10,12)) +
  scale_y_continuous(breaks = c(0,2,4,6,8,10)) +
  geom_smooth(method = "lm", formula = (y ~ exp(-x)), se = FALSE, color = "black") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 20, color = "black"),
        axis.title = element_text(size = 24),
        text = element_text(family = "Times"))


sol_summary <- solubilization_rates %>% 
  ungroup() %>% 
  summarize(mean_SR = mean(daily_solubilization),
            minSR = min(daily_solubilization),
            maxSR = max(daily_solubilization),
            sdSR = sd(daily_solubilization))


solubilization_chl_C <- lm(log(daily_solubilization) ~ Chl_C, data = solubilization_rates)

summary(solubilization_chl_C)

prediction_interval <- exp(predict(
  solubilization_chl_C, 
  newdata = solubilization_rates,
  interval="prediction",
  level = 0.95
  ))


```

# Stats: Exponential Regression of Solubilization Rates and Chl:C

```{r}

solub_chlC_lm <- lm(daily_solubilization ~ exp(-Chl_C) ,data =solubilization_rates)

summary(solub_chlC_lm)

Chl_C <- solubilization_rates$Chl_C

predicted_solubilization_rates <- predict(solub_chlC_lm, solubilization_rates)

new_x      <- seq(min(Chl_C), max(Chl_C), )

prediction <- exp(predict(solub_chlC_lm, newdata = list(x = new_x)))

exp_line   <- data.frame(x = new_x, y = prediction)



```



# Senescent DOC vs NPP 

```{r}


doc_NPP_senescent <- senescent_kelp %>% 
  ggplot(aes(x = NPP_umolC_g_hr, y = DOC_umolC_g_hr, color = PAR)) +
  geom_point(size = 7) +
   geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 2) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(NPP~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(color = "PAR") + 
  scale_x_continuous(limits = c(-50,300),breaks = c(-50,0,100,200,300)) +
   scale_color_viridis_b(breaks = c(0,50,100,250,500,1000,1500), limits = c(0,1500), name = expression(paste("PAR (µmol ",m^-2, s^-1,")"))) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))



doc_PAR_senescent <- senescent_kelp %>% 
  ggplot(aes(x = PAR, y = DOC_umolC_g_hr)) +
  geom_point(size = 7) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  xlab(expression(expression(paste("PAR (µmol ",m^-2, s^-1,")")))) +
  labs(color = "PAR") + 
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 24),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 25),
        text = element_text(family = "Times"))


# Model II Senescent DOC vs NPP and PAR

senescent_DOC_NPP <- lmodel2(DOC_umolC_g_hr~NPP_umolC_g_hr, data = senescent_kelp)


senescent_DOC_PAR <- lmodel2(DOC_umolC_g_hr~PAR, data = senescent_kelp)


```




# Modeling Kelp Biomass Loss from Senescence into POC and DOC
## Using solubilization rate of 0.025
## Using POC shedding rate of 0.0034 (+ 2SD from Average POC shedding rate)


```{r}

kelp_biomass <- 100

senescence_function <- function(x){
  biomass_proportion <- (100*(1 -(-0.0002*0.0036 + 0.0000058*(0.0036^2) + 0.0036)*x))
  print(biomass_proportion)
}

biomass_remianing <- sapply(51:100, senescence_function)





age <- seq(51,100, by = 1)
timepoint <- seq(1,50, by = 1)


age_df <- data.frame(age, biomass_remianing)

biomass_solubilization_plot <- age_df %>% 
  ggplot() +
  geom_line(aes(x = age, y = biomass_remianing, color = "black"),size = 2) +
  geom_line(aes(x = age, y = 100-biomass_remianing, color = "red"), size = 2) +
  ylab("% Biomass") +
  xlab("Age (d)") +
  labs(color = c("Biomass", "DOC")) +
  scale_color_manual(values = c("black", "red"), labels = c("Biomass", "DOC")) +
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_blank(),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))


```


# Stoichiometry of DOM 

```{r}

MP_age_data_sum %>% 
  group_by(age_d) %>% 
  summarize(age = age_d, mean_DOC = mean(delta_DOC_uMC), mean_DON = mean(delta_DON_uMN), sd_DOC = sd(delta_DOC_uMC), sd_DON = sd(delta_DON_uMN)) %>% 
  ggplot(aes(x = age, y = mean_DOC/mean_DON)) +
  geom_point(size = 5) +
  labs(x = "Age (d)",
       y = "Exudate C:N") +
  theme(panel.background = element_rect(fill = "white", color = "black", size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 24),
        legend.key.height = unit(0.5, "inch"),
        legend.key.width = unit(0.4, "inch"),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))





```

# Bell & Siegel Chl a vs Age Data 

```{r}


bell_chla_age <- read_xlsx("Cohort_Chl.xlsx") %>% 
  filter(prop_original <= 1)


bell_prop_chla_age_p <- bell_chla_age %>% 
  ggplot(aes(y = Age, 
             x = prop_original)) +
  geom_point(size = 6) +
  xlab(expression(Chl~a~(nmol ~ g^{"-1"}))) +
  labs(y = "Age (d)") + 
  scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))


lm_age_proportion_initial_chlA <- lm(Age~prop_original, data = bell_chla_age)

lm_age_proportion_initial_chlA
summary(lm_age_proportion_initial_chlA)

```

# Normalize Chl values to orignal 

```{r}


MP_modeled_age_df <- MP_DOC_NPP_phys_env_sum_noflags %>% 
  filter(season != "NA",
          DOC_umolC_g_hr >0) %>% 
  select(season, age_d, chlA_nmol_g, DOC_umolC_g_hr, tissue_C_percent, PAR) %>% 
  group_by(season) %>%
  summarise(norm_chlA = (chlA_nmol_g/mean(chlA_nmol_g[age_d == min(age_d)])),
            DOC_umolC_g_hr = DOC_umolC_g_hr,
            age_d = age_d,
            norm_chlA = norm_chlA,
            modeled_age_d = norm_chlA*(-76.185) + 100.468,
            daily_solubilization = ((DOC_umolC_g_hr/1000/1000*12)*24)/((tissue_C_percent/100))*100,
            PAR = PAR) # 


model_age_DOC_lm <- lm(DOC_umolC_g_hr ~ poly(modeled_age_d,2, raw = TRUE) , data = MP_modeled_age_df)

summary(model_age_DOC_lm)


model_age_solubilization_lm <- lm(daily_solubilization ~ poly(modeled_age_d,2, raw = TRUE) , data = MP_modeled_age_df)

summary(model_age_solubilization_lm)


doc_production_chlA_normalized <- MP_modeled_age_df %>% 
  ggplot(aes(x = norm_chlA*100, 
             y = DOC_umolC_g_hr)) +
  geom_point(aes(color = season), size = 3) +
  scale_color_manual(values = c("blue", "black")) +
  ylab(expression(DOC~Production~(µmolC ~ g[DW]^{"-1"} ~ hr^{"-1"}))) +
  labs(x = "Normalized ChlA (% Initial)",
       color = "") +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75,100)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        legend.position = c(0.8,0.8),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))
  
  
age_modeled_age <- MP_DOC_NPP_phys_env_sum_noflags %>% 
  filter(season != "NA") %>% 
  select(season, age_d, chlA_nmol_g, DOC_umolC_g_hr) %>% 
  group_by(season) %>%
  summarise(norm_chlA = (chlA_nmol_g/mean(chlA_nmol_g[age_d == min(age_d)])),
            DOC_umolC_g_hr = DOC_umolC_g_hr,
            age_d = age_d,
            modeled_age_d = norm_chlA*(-62.266) + 95.685) %>% 
  ggplot(aes(y = age_d, 
             x = modeled_age_d)) + 
  geom_point(size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  scale_x_continuous(breaks = c(0,25,50,75,100), limits = c(0,100)) +
  scale_y_continuous(breaks = c(0,25,50,75,100), limits = c(0,100)) +
  theme(panel.background = element_rect(fill = "white", 
                                        color = "black", 
                                        size = 1.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(1,1,1,1, "cm"),
        strip.background = element_rect(fill = "white", 
                                        color = "black"),
        legend.key = element_rect(fill = "white"),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20),
        axis.text = element_text(size = 30, 
                                 color = "black"),
        axis.title = element_text(size = 30),
        text = element_text(family = "Times"))






```

